\chapter{Exchangeable Array Processes}

\begin{center}
    \textbf{Abstract}
\end{center}

\red{todo}

% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The goal of this paper is to give a framework for building large combinatorial objects in a ``consistent'' way and to attempt to extend the work of \cite{aldous2010_more_uses_of_exchangeability} which sought, among other aims, to develop a general theory of limits of combinatorial objects.

\red{say more about why this may be useful/interesting, and the relationship between exchangeability and sampling.}


\indent Given a transient Markov chain $\Xn$ with a countable state space $S$ along with a base state $x_0 \in S$ from which the chain is irreducible, some of the goals of Martin boundary theory are:

\begin{enumerate}[label=(\roman*)]
    \item Complete $S$ into a compact metric space by adding ``points at $\infty$'' to which the Markov chain is converging. 
    %
    \item Characterize all the nonnegative/bounded harmonic functions $h :S \to [0,\infty)$.
    %
    \item Develop a notion of conditioning the chain ``at infinity'' and understanding the behavior of these conditioned chains.
\end{enumerate}

\noindent Since the chain $X_n$ is transient and discrete, it has no natural notion of convergence within $S$. Endowing it with a metric structure 
allows us to discuss convergence in a more topological sense, and compactness gives us 
the suitable regularity to ensure such convergence often exists. A review of the Doob-Martin boundary can be found in Appendix~\ref{sec:doob_martin_background_appendix}.

% Exchangeable Array Processes %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exchangeable Array Processes}

The following details can be found in \cite{Austin_arrays} or Chapter 7 of \cite{KallenbergProbabilisticSymmetries}.

Given a set $A$ let $A^{(d)}$ denote the set of all subsets of size $d$ of $A$.
 A \textit{d-dimensional (jointly) exchangeable array} is a collection of random variables 
 $(X_e)_{e \in \N^{(d)}}$ indexed by $d-subsets$ of $\N$ whose distribution is invariant under 
 simultaneous permutation of the coordinates. Namely, letting $Sym(\N)$ be the set of finite 
 permutations of $\N$ then
%
\[(X_{\sigma(e)})_{e \in \N^{(d)}} \overset{\text{law}}{=} (X_e)_{e \in \N^{(d)}}, \quad \forall \sigma \in \mathrm{Sym}(\N),\]
%
\noindent where $\sigma(e) = (\sigma(i_1), ..., \sigma(i_d))$ for $e = (i_1,..., i_d)$. 

If $(X_e)_{e \in \N^{(d)}}$ is an exchangeable array there are two natural Markov processes associated to it: a ``labeled'' chain $L_n$ and an ``unlabeled'' one $W_n$. Informally, the labeled chain $L_n$ is produced by just revealing the entries with maximal coordinate $n$ at step $n$. Namely, the finite labeled chain $L_n$ is just the ``top left corner'' of the array: the restriction of the array to entries corresponding $[n]^{(d)}$. In this framework, the full array $(X_e)$ plays the role of the projective limit of the labeled chain which often appears in Doob Martin theory. 

The unlabeled chain $M_n$ is produced from the labeled one by ``forgetting the order'' of the rows, columns, et cetera and just keeping track of the set of values appearing in each row, column, and so on. More formally:

\begin{definition}[Exchangeable Array Processes]
Given an exchangeable array $(X_e)_{e \in \N^{(d)}}$ with entries in some Borel space $S$, define a graded Markov chain $L_n$ whose state space is $E := \bigcup_{n \in \N} E_n$, where $E_n := S^{[n]^{(d)}}$ by:
%
    \[(L_n)_e := X_e\quad \forall e \in [n]^{(d)}.\]
%
\noindent Define an equivalence relation $\sim$ on $E_n$ by saying $A \sim B$ if there exists some $\sigma \in Sym([n])$ so that $A_e = B_{\sigma(e)}$ for all $e \in [n]^{(d)}$, where $\sigma(e) = (\sigma(i_1), \sigma(i_2), ..., \sigma(i_k))$ just permutes all elements of the subset $e$ by the same permutation. With this, define the unlabeled chain:
%
\[M_n := [L_n]_\sim.\]
%
\noindent a Markov chain on the space of equivalence classes of finite $d-dimensional$ arrays.
\label{def:exchangeable_array_process}
\end{definition}
%
\noindent In this paper, we will almost exclusively think of $S$ as countable, and often finite. Note that $L_n$ is trivially a Markov process in the sense that any sequence of random variables becomes a Markov process if the state space is just the entire history of the process. In general, projections of Markov chains are not Markovian, but in this case, exchangeability ensures that the Markov property is preserved.

% lemmma
\begin{lemma}[Markov Property of the Unlabeled Chain]
    When the state space $S$ is countable, the process $M_n$ is a Markov chain. 

    \label{lem:index_equivalence}
\end{lemma}
%



\noindent The above lemma is true for any initial distribution, but we are almost exclusively interested in the case $M_0$ is the empty array.  It is likely possible to extend this framework to more general equivalence relations, possibly under some constraints on the initial conditions.
%
\noindent The main benefit of this framework is the following theorem that shows all exchangeable arrays have a representation through sampling from some hypercube. 


\begin{theorem}[Representation Theorem]
\noindent An exchangeable array on $d$-subsets of $\N$ has a representation of the form:

\[\{X_e\}_{e \in \N^{(d)}} =^d \left\{f\left(U_\emptyset, (U_i)_{i \in e}, (U_a)_{a \in {e \choose 2}}, ..., (U_a)_{a \in {e \choose d-1}}, U_e)\right) \right\}_{e \in \N^{(d)}},\]

\noindent for some middle symmetric $f:[0,1]^{2^d} \to S$ and iid uniforms. Moreover, any array of this form is exchangeable.

\label{thm:representation_theorem_of_exchangeable_arrays}
\end{theorem}

\noindent A middle-symmetric function satisfy the following for any $\sigma \in Sym([d])$:
%
    \[f\left(x_\emptyset, (x_i)_{i \in [d]}, (x_a)_{a \in [d]^{(2)}}, ..., x_{[d]})\right) = f\left(x_\emptyset, (x_{\sigma(i)})_{i \in [d]}, (x_{\sigma(a)})_{a \in [d]^{(2)}}, ...,  x_{[d]})\right).\]
%
In words, you are allowed to apply the same permutation to the indices ``at each level'' separately. 

Throughout this paper, we will encounter many domains of the form $[0,1]^{2^k}$ for some $k$. For ease of notation, we will assume the coordinates are always indexed by subsets of $[k]$ as in the definition of a middle symmetric function above.

In the case of one dimensional arrays, this representation theorem reduces to the standard De Finetti's theorem which says an exchangeable sequence is a mixture of iid sequences and thus takes the form $(X_n) =^d (f(U, U_n))$. Moreover, two-dimensional arrays take the form $f(U, U_i, U_j, U_{ij})$ for some function $f$ with $f(w,x,y,z) = f(w,y,x,z)$. 

As \cite{aldous2010_more_uses_of_exchangeability} points out, this quantifies the fundamental relationship between 
exchangeability and sampling that is already so well known: if we can get a hold of the representing function 
$f$, it gives us a simple procedure for generating the corresponding combinatorial objects.
Next, we now point out a simple way of generating new exchangeable arrays from old ones.

\begin{lemma}[State-Space Projections of Exchangeable Arrays]
    If $(X_e)$ is a exchangeable array with entries in $S$ and $g:S \to S'$ is measurable then $(g(X_e))$ is an exchangeable array of the same dimension. Moreover, if $f$ is the representing function of $(X_e)$, then $g \circ f$ is the representing function of $(g(X_e))$.

    \label{lem:state_equivalence}
\end{lemma}

\noindent Here we are essentially defining an equivalence relation on the entries of the array where we let $s \sim s'$ if $g(s) = g(s')$. 
We call this the \textit{state space equivalence}. 

Combined with definition \ref{def:exchangeable_array_process}, this shows that there are essentially three layers of Markov processes associated to a given exchangeable array: (i) the initial ``labeled'' exchangeable array process corresponding to ($X_e$), then (ii) the ``labeled'' array process corresponding to $(g(X_e))$ where we apply an equivalence to the state space and lastly (iii) the ``unlabeled'' chain where we apply an equivalence relation to the coordinates of the array. See the example below on ballot sequences for an example where all three layers are utilized.


\begin{example}[Polya's Urn]
\label{ex:polyas_urn}
    In the case $d=1$, an exchangeable array is just an exchangeable sequence. It is 
    well-known the sequence of balls drawn from a Polya's urn is an exchangeable sequence. 
    In this case, we can think of $X_i$ of taking values in $[m]$ if we have $m$ colors. 
    The labeled chain $L_n$ is the sequence of balls, and the unlabeled chain $W_n$ just
     keeps track of the counts of balls of each color. This has representation function

    \[f(x,y) = \sum_{i=1}^m i \cdot \ind \left\{y \in [t_{i-1}, t_{i})\right\} \qquad t_i := \sum_{k=1}^i p_k(x),\]

    \noindent where $(p_1(U), ..., p_m(U)) \sim Dirchlet(\alpha_1,..., \alpha_m)$ for $\alpha_i$ initial balls of color $i$. 
\end{example}

\begin{example}[Graphon Chain]
    Given any graphon $w$ there is a sequence of finite graphs $G_n$ converging to $w$ in the
    cut metric (equivalently the Doob-Martin topology by \cite{Grubel_2015_graphons}) 
    constructed via sampling $U_i \sim Unif[0,1]$ and then connecting vertex $i$ to vertex
     $j$ with probability $w(U_i, U_j)$. Thus a graphon can be represented as an ``infinite adjacency matrix'' 
     which forms an exchangeable array with entries
    %
    \[X_e = \ind \{U_e \leq w(U_i, U_j)\} \quad \text{ for } e = (i,j) \in {\N \choose 2},\]
    %
    \noindent Here the representation function $f(x_\emptyset, x_i, x_j, x_{ij})$ does not depend on $x_\emptyset$.
    \label{example:graphon_chain}
\end{example}

\begin{example}[d-Uniform Hypergraph Chain]
    We can extend the graphon example to d-uniform hypergraphs, which have hyperedges between subsets of $[n]$ of size $d$, initially explored in \cite{hypergraphons}. Namely, define a \textit{hypergraphon} as a middle-symmetric function $W:[0,1]^{2^d-2} \to [0,1]$ with coordinates indexed by proper, nonempty subsets of $[d]$. Then, analogous to graphons, this has representation function:

    \[f\left(x_\emptyset, (x_i)_{i \in [d]}, ..., x_{[d]})\right) = \ind \Big \{x_{[d]} \leq W\left((x_i)_{i \in [d]}, ...,  (x_a)_{a \in [d]^{(d-1)}}\right) \Big \}.\]
    
    \noindent As we will see in Corollary~\ref{cor:classification_binary_array_processes} below, these are in a sense the only binary array processes. 
\end{example}

\begin{example}[Ballot Sequences]
Given some total ordering of $I_0 := \bigcup_{i \in \N} \{a_i, b_i\}$ so $a_i < b_i$ for each $i$, there 
are are 6 possible ``interweavings'' of the letters:
%
\[a_ib_ia_jb_j \quad a_i a_j b_i b_j \quad a_i a_j b_j b_i \quad a_j a_i b_i b_j \quad a_j a_i b_j b_i \quad a_j b_j a_i b_i.\]

\noindent This ordering can be encoded by a 2-dimensional array $X_{ij}$ with values in $\{1,...,6\}$ 
describing which of these relationships the pairs $(a_i, b_i), (a_j, b_j)$ share. 
If we equate the interweavings above which have the $a's$ and $b's$ in the same locations, 
the unlabeled chain generates \textit{ballot sequences}: strings of $a's$ and $b's$ where every prefix
has at least as many $a's$ as $b's$. In~\cite{ChoiEtAlGrowingBallotSequences}, the boundary is characterized through
a pair of diffuse marginals $(\zeta, \mu)$ encoding the relative density of $a's$ and $b's$ in the limit. 
\end{example}

\noindent Exchangeable array processes also arise naturally in the the theory of 
exchangeable partitions through Kingman's paintbox theorem \cite{Austin_arrays}, random permutations through
permutons~\cite{Grubel_2023_permutons}, planar binary trees through Remy's chain \cite{DM_Remy}, and growing 
random strings~\cite{choi2016doobmartincompactificationmarkovchain}.

These and the preceeding examples give a common strategy for embedding combinatorial processes as random arrays.
If a random structure consists of certain ``parts'' (e.g. vertices of a graph) and is completely determined by the 
relationships within each subset of $d$ parts (e.g. which vertices are connected by an edge), then it can be 
encoded as a d-dimensional array whose entries encode the relationship. Exchangeability then enforces the
constraint that the distribution of the structure is invariant under relabeling of the parts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Doob-Martin Boundary Theory for Exchangeable array Processes %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Doob-Martin Boundary of Exchangeable Array Processes}

\noindent The backwards dynamics of the labeled chain are deterministic: given $L_n$ you simply restrict 
to $e \in [n-1]^{(d)}$ to obtain $L_{n-1}$. Thus, its Doob-Martin boundary is almost trivial.
%
\begin{lemma}[Boundary of the Labeled Chain]
    The boundary of $L_n$ consists of the deterministic arrays $x \in \mathrm{supp}(X_e)$.
    \label{lem:boundary_labeled_chain}
\end{lemma}

\noindent Next, we discuss how state-space equivalence and unlabeling affect this boundary.
%%% state equivalence
\subsection{State Equivalence}

\noindent As discussed earlier, given the exchangeable array $(X_e)$ we can generate a new exchangeable array $(g(X_e))$, and by lemma \ref{lem:boundary_labeled_chain} the boundary of the corresponding labeled array processes are just the deterministic arrays on the relevant state spaces.

\zack{We know from the previous section the boundary of $g(X_e)$ is again just the deterministic arrays $(y_e)$. But I want to relate this to the boundary of $X_e$ in a nice way. The preimage $g^{-1}((y_e))$ of a deterministic array includes random arrays $(X_e)$ so that $supp(X_e) \subseteq g^{-1}(y_e)$. Thus we get a whole collection of bridges of $(X_e)$, some of which are extremal and some of which are mixtures of these extremal ones. I should go back to the ballot chain example to see how to possibly interpret this as a measure on the boundary of the labeled chain. }


%%% Labeled Chain
\subsection{Boundary of the Unlabeled Chain}

\noindent Now we seek to understand the boundary of the unlabeled chain $M_n$ obtained via our equivalence $\sim$. Recall that there is a correspondence between infinite bridges of a given Markov chain and its nonnegative harmonic functions with $h(x_0) = 1$ via the Doob h-transform. 

\begin{lemma}[Exchangeable Bridges of the Labeled Chain]
    Any infinite bridge of the labeled exchangeable array process is an array process, and it is exchangeable if and only if the corresponding harmonic function $h$ is constant on the equivalence classes of $\sim$. 

    \label{lem:harmonic_functions_labeled}
\end{lemma}



An immediate consequence of this is that the harmonic functions of the unlabeled chain are in direct correspondence with the bridges of labeled chain which come from exchangeable arrays.

\begin{corollary}[Bridges of the Unlabeled Chain]
    If $M^{\tilde{h}}_n$ is an infinite bridge of the unlabeled chain $M_n$, then $M^{\tilde{h}}_n$ equals the unlabeled process associated to some exchangeable array $(X_e^h)$. In particular, if $\tilde{h}$ is the harmonic corresponding to the bridge $M^{\tilde{h}}_n$, then $h(x) := \tilde{h}([x])$ is the harmonic corresponding to the array $(X_e^h)$.

    \label{cor:unlabeled_associated_to_exchangeable_array}
\end{corollary}

\daniel{Should this corollary include a statement saying how $(X_e^h)$ and the array $(X_e)$ generating $M_n$ relate?}



The relation $\sim$ extends to infinite $d-dimensional$ infinite arrays $x$ in the obvious way: $x \sim y$ if there is $\sigma \in Sym(\N)$ so $x = \sigma(y)$. We say a measurable collection $I$ of arrays is \textit{$\nu-invariant$} if permutations do not change up up to $\nu-null$ sets:
\[ I \overset{\nu-a.s.}{=} \tau I \quad  \quad \forall \tau \in Sym(\N).\]


We say an exchangeable array with distribution $\nu$ is \textit{ergodic} if $\nu(I) \in \{0,1\}$ for all $\nu-$invariant sets $I$. 

 We say an exchangeable array is \textit{dissociated} if for any disjoint index sets $I_1,...,I_m \subset \N$ the arrays $\{X^i_e\}_{e \in I_i^{(d)}}$ are independent. For example, the exchangeable array corresponding to a graphon is disassociated because in this setting, this condition just asserts that the random  graphs corresponding to disjoint collections of vertices are independent in the graphon model, which follows immediately from the fact that the presence of edges are determined independently.

 The following theorem shows that all these aforementioned properties characterize the Doob-Martin boundary of the unlabeled chain.

\begin{theorem}[Extremal Bridges and Ergodic Arrays]
    For any infinite bridge of the unlabeled exchangeable array process $M_n$, the following are equivalent:

    \begin{enumerate}
        \item The infinite bridge is extremal.
        \item Its corresponding exchangeable array is ergodic.
        \item Its corresponding exchangeable array is dissociated.
        \item Its corresponding representation function $f$ does not depend on $U_\emptyset$ (almost surely). 
    \end{enumerate}
    \label{thm:boundary_is_ergodic_arrays}
\end{theorem}


%


This also shows that the representation function of a given unlabeled bridge can in a sense be written as a mixture of the extremal representation functions $f^{\tilde{K}(\cdot,\alpha)}$, which do not depend on the first coordinate by the previous theorem. More specifically, if we let $g: [0,1] \to \overline{E}$ be a function so that $g(U)$ has distribution $\mu_{\tilde{h}}$ this shows that $f^{\tilde{h}}(U, (U_i), ..., U_e) = f^{g(U)}((U_i), ..., U_e)$ and so our representation function is a ``mixture'' of the extremal representation functions.


\begin{remark}
    Our observation in \ref{example:graphon_chain} that the representation function $f(q,x,y,u) := \ind \{u \leq w(x,y)\}$ for our graphon $w$ does not depend on $q$ is then, according to the previous lemma, just the statement that graphons correspond to the extremal infinite bridges, and hence the minimal Doob-Martin boundary, of the graphon chain. \red{In fact I think for binary arrays the extremal representation functions will always look like $\ind\{u \leq f((x_i), (x_{ij}), ...)$ up to the Lebesgue isomorphisms in the uniqueness theorem. So the binary exchangeable array processes look exactly like the graphon chain but with edges between $d-subets$ of the graph! So like a hypergraph chain but where the edge degree is fixed to be $d$. So the hypergraph chain is just like an infinite collection of exchangeable array chains, one for each $d$. }
\end{remark}

\daniel{Can we then say that the full boundary (not just minimal) comes from stochastic block models then?}

\noindent \red{Another interpretation of the above result is that the boundary of the unlabeled chain is in terms of measures on the labeled boundary.}


% Embedding Density of Representation Functions %%%%%%%%%%%%%%%%%%%%
\newpage
\section{Embedding Density of the Representation Functions}

\noindent The of embedding a smaller structure into a larger structure plays a pivotal role in the theory of combinatorial processes. \red{give a little more motivation}.

Two of our main examples, graphons and permutons, already have natural notions of embeddings in their pre-existing theory. The aim of this section is to try to fit these notions of embeddings into the framework of exchangeable arrays, in particular binary arrays, expanding upon the work of \cite{borgs2007convergent}. First, we introduce some notation: for a finite binary array of d-dimensional array $F$ of size $k$ let:
%
\[\mathbf 1(F) := \{ S \in [k]^{(d)} :  F_S = 1 \}.\]
%
\noindent denote the collection of entries of our array which are one. Moreover, given a vector $x := (x_e)_{e \subseteq [n]} \in [0,1]^{2^n}$ and a subset $A \in [n]^{(d)}$ denote:
%
\[x_{(A)} := (x_\emptyset, (x_i)_{i \in A}, (x_{ij})_{ij \in A^{(2)}}, \;... \;, x_A) \in [0,1]^{2^d},\]
%
\noindent analogous to how the representation function $f$ is indexed in Theorem \ref{thm:representation_theorem_of_exchangeable_arrays}. Namely, if $x$ is indexed by all subsets of $[n]$ then $x_{(A)}$ restricts to only subsets of $A$. Lastly, for a finite $d-dimensional$ array of size $k$ let $|x| := k$ denote its size.

\subsection{Embeddings of Labeled States}

Suppose we have arrays $(x_e)_{e \in [n]^{(d)}}$ and $(y_e)_{e \in [n+m]^{(d)}}$. Then for the labeled chain we say \textit{x is embedded in y} if $y_e = x_e$ for all $e \in [n]^{(d)}$ and denote this by \red{$x \leq y$}. Namely, we obtain $x_e$ by looking at the ``top left corner'' of the array $y_e$. From the viewpoint of our Markov chain, this just means that the array $(y_e)$ is reachable starting in state $(x_e)$. This is an extension of the standard notion of an induced subgraph in the theory of graphs, and in fact makes sense beyond just binary arrays.

Likewise, the standard notion of graph embedding extends to the following notion for binary arrays: we say a binary array $x$ is \textit{weakly embedded} in binary array $y$, denoting this by $\red{x \lesssim y}$, if $y_e = 1$ whenever $x_e = 1$. This notion of embedding is the one which the theory of graphons is built on, but as we show below from a topological perspective the two notions are equivalent. In a sense, the former notion of embedding is more natural for exchangeable arrays because it makes sense for nonbinary arrays.


\subsection{Embeddings of Unlabeled States}

\noindent Again suppose we have arrays $(x_e)_{e \in [n]^{(d)}}$ and $(y_e)_{e \in [n+m]^{(d)}}$. For the unlabeled chain, we want to incorporate the idea that the arrays are only defined up to a permutation of the indices. Thus say $[x]_\sim$ is (weakly) embedded in $[y]_\sim$ if there exists $\sigma \in Sym([n+m])$ so $(x_e)$ is (weakly) embedded in $(y_{\sigma(e)})$ in the sense of our labeled chain. Equivalently, if there exists an injective function $\phi:[n] \to [n+m]$ so $(x_{\phi(e)}) = (y_{e})$. 

In this case, there can be more than one embedding if there are several different choices of $\sigma$ that produce an embedding. For example, if both $x,y$ are arrays of all zeros of their respective sizes then any $\sigma \in Sym([n+m])$ will produce an embedding. This suggests a notion of \textit{embedding density} analogous to the theory of graphons and permutons.
%
\begin{align}
    emb(x,y) & := \frac{| \{\sigma \in Sym([n+m]) \; : \; (x_e) \leq (y_{\sigma(e)}) \}|}{(n+m)!} \notag\\[7pt]
            & = \frac{| \{\phi:[n] \xrightarrow{injective} [n+m] \; : \; (x_{\phi(e)}) \leq (y_e) \}|}{(n+m) (n+m-1) \cdot ... \cdot (n+1)} 
    \label{def:embedding_density}
\end{align}

\noindent for $x \in E_n, y \in E_{n+m}$. Note that the embedding density is constant on each equivalence class of $x$ so this is well-defined as a function on these equivalence classes.  Likewise we can define the \textit{morphism density} by removing the injectivity constraint:
%
\begin{equation}
    mor(x,y) := \frac{| \{\phi:[n] \to [n+m] \; : \; (x_{e}) \leq (y_{\phi(e)}) \}|}{(n+m)^n}.
\end{equation}

\noindent Analogously, for binary arrays we can define the \textit{weak embedding/morphism density} $emb^W$ and $mor^W$ respectively for by replacing $\leq$ by $\lesssim$. These share the following relationships:

\begin{lemma}[Embedding and Morphism Densities]
    For binary arrays $x,y$ let $x'$ be an array of the same size as $x$. Then:

    \begin{enumerate}[label=(\roman*)]
        \item $emb(x,y) \leq emb^W(x,y)$. 
        \item $emb^W(x,y) = \sum_{x' \geq x} emb(x,y)$
        \item $emb(x,y) = \sum_{x' \geq x} (-1)^{|1_x| - |1_{x'}|} \cdot emb^W(x',y)$

        \item $|mor^W(x,y) - emb^W(x,y)| \leq \frac{1}{|y|} \cdot {\;|x|\; \choose 2}$ 
        
        \item  $|mor(x,y) - emb(x,y)| \leq \frac{1}{|y|} \cdot {\;|x|\; \choose 2}$
    \end{enumerate}   

    \label{lem:relating_different_notions_of_embedding}
\end{lemma}



\noindent The above in fact shows that from a topological perspective, all these notions of embedding are equivalent. Thus in formalizing a notion of convergence, either definition will suffice. Moreover, as the above observations may suggest, there is a relationship between such processes and the backwards exchangeable processes of \red{the other paper}.

\begin{lemma}[Doob-Martin Convergence and Embedding Densities]
    The unlabeled chain $M_n$ of an exchangeable array process is uniformly backwards exchangeable with labels $\sigma \in Sym([n])$. Moreover, the following are equivalent:
        \begin{enumerate}[label=(\roman*)]
            \item The sequence $y_N$ converges in the Doob-Martin topology.
            \item The embedding density $emb(x, y_N)$ converges for all $x$.
            \item The morphism density $mor(x, y_N)$ converges for all $x$. 
        \end{enumerate}

    \noindent Furthermore, for binary arrays convergence of the weak embedding/morphism density is also equivalent.

    \label{lem:equivalence_of_dm_and_density_convergence}
\end{lemma}



\zack{We would like to say something stronger; that every backwards exchangeable chain, subject possibly to some constraints, can be encoded in such an array. But I think we are limited here to only chains that can be labeled by a single permutation, rather than our sort of more arbitrary labeling scheme discussed at the start of the paper.}\\



\zack{The key idea for the below lemmas/proofs is that most of the fundamental results about the embedding densities only interact with graphs through bounding the number of injective maps $\phi : [k] \to [n]$ that appear in equation \eqref{def:embedding_density}. Hence these results extend almost verbatim to our more general results on exchangeable arrays, altough I suspect the concentration bounds could be improved to depend on the dimension $d$ of the array.}

\noindent In the case of our graphon and permuton chains, the notion of embedding can be extended to the limiting object, graphons and permutons, via sampling. Namely, for a graphon $w$ its embedding density $t(\cdot, w)$ is essentially defined so that for each $n$ and finite graph $F$:
%
\[t(F, w) = \E t(F, G_n(w)),\]

\noindent where $G_n(w)$ is the random graph generated by sampling $n$ uniforms $U_1,..., U_n$ and connecting vertices $i,j$ with probability $w(U_i, U_j)$.

Analogously, we would like to extend this notion of an embedding density to our limiting objects: the representation functions $f$ of our exchangeable arrays. This suggests the definitions:
%
\begin{equation}
    emb(x, f) := \E \; emb(x, X_n(f)), \qquad  emb^W(x, f) := \E \; emb^W(x, X_n(f)),
\end{equation}

\noindent where $X_n(f)$ is the array indexed by $[n]^{(d)}$ generated by:
%
\[(X_n(f))_e := f(U_\emptyset, (U_i)_{i \in e},..., U_e).\]

\noindent We will now show these expectations are independent of $n$ and have an integral representation analogous to the embedding density of graphons. The key insight is that many of the preliminary results in the theory of graphons rely purely on controlling the embedding densities by bounding the number of injective functions $\phi:[k] \to [n]$, and thus generalize immediately to exchangeable array processes.

\begin{lemma}[Embedding Densities of Representation Functions]
If $F$ is a size $k$ array $d-dimensional$ array an $f$ the representing function of an exchangeable array $X$ with $ k \geq d$ then for \red{$n \geq k$}:

\begin{equation}
    emb^W(F,f) := \E \; emb^W(F, X_n(f)) = \int_{[0,1]^{2^k}}  \prod_{S \in \ind(F)} f(x_{(S)}) \; d\lambda_{2^k} 
    \label{eqn:integral_formula_for_embedding_density_of_representing_function}
\end{equation}
\begin{equation}
    emb(F,f) := \E \; emb(F, X_n(f)) = \int_{[0,1]^{2^k}}  \prod_{S \in [k]^{(d)}} \ind \{f(x_{(S)}) = F_S\} \; d\lambda_{2^k}.
\end{equation}
%
\noindent In particular, it is independent of $n$. Moreover we have concentration:
%
\begin{align*}
    \P\left( |emb^W(F, X_n(f)) - emb^W(F, f)| > \epsilon \right) &  \leq 2 \exp \left(-\frac{\epsilon^2}{\red{2}k^2} n \right)\\
    % 
	    \P\left( |emb(F, X_n(f)) - emb(F, f)| > \epsilon \right) & \leq 2 \exp \left(-\frac{\epsilon^2    }{\red{2}k^2} n \right).
\end{align*}
%
\noindent Thus,
\[
mor(F, X_n(f)), emb(F, X_n(f)) \xrightarrow{a.s.} emb(F,f)
\]
\noindent and
\[
mor^W(F, X_n(f)), emb^W(F,X_n(f)) \xrightarrow{a.s.} emb^W(F,f).
\]
\noindent In particular, the results of Lemma \ref{lem:relating_different_notions_of_embedding} (i) - (iii) extend to $emb^W(F,f)$ and $emb(F,f)$.


\label{lem:integral_representations_of_embedding_densities}
\end{lemma}

\noindent This is perfectly analogous to the theory of graphons which shows:
%
\[\E (emb(F, G_n(w)) = \int_{[0,1]^k} \prod_{(i,j) \in E_F} w(x_i, x_j) \; dx.\]
%
\noindent To see why, note that in the case $d=2$:
%
\[ \int_{[0,1]^{2^k}}  \prod_{S \in \ind(F)} f(x_{(S)}) \; d\lambda_{2^k} = \int_{[0,1]^{2^k}}  \prod_{(i,j) \in \ind(F)} f(x_\emptyset, x_i, x_j, x_{ij}) \; d\lambda_{2^k}. \]
%
\noindent Thus the integrand depends only on the $1 + k + {k \choose 2}$ coordinates $x_e$ for subset $e$ of $[k]$ of size at most two. For graphons the terms inside the product become $\ind\{x_{ij} \leq w(x_i, x_j)\}$ so we can integrate out the coordinate $x_\emptyset$ and the coordinates $x_{ij}$ (which only appear in one term of the product) to get the familiar formula from graphons. 

In general, we can see that the integral in \eqref{eqn:integral_formula_for_embedding_density_of_representing_function} only relies on the $\sum_{i=0}^d {k \choose d}$ coordinates corresponding to subsets $e \subseteq k$ of size at most $d$, so many of the coordinates in the integral are superfluous. We leave them in for simplicity of notation.  


% Metrizing Convergence of Exchangeable Array Processes %%%%%%%%%%%%%%
\newpage
\section{Metrizing Convergence of Exchangeable Array Proceses}

 The paper \cite{Grubel_2015_graphons} showed graphon convergence and its associated topology is equivalent to the one given by the Doob-Martin theory. While the Doob-Martin topology is trivially metrizable as it is essentially just a Euclidean space, this metric is sort of ad hoc and not typically useful. The real success of the theory of graphons was the introduction of a metric, the cut metric, which is reasonably interpretable and proves to be a valuable tool for explicit calculation of many interesting properties of their associated graphs.

 \red{Firstly we will develop a metric for binary arrays and latter we will discuss how it can be extended to any exchangeable array process.} As discussed above, d-dimensional binary arrays are essentially just d-uniform hypergraph chains. Thus we can view this first section as extending the cut metric to d-uniform hypergraphs and hypergraphons. 
 
 We saw in the above example that the graphon $w$ of a graphon chain is closely related to the representation function $f$ of its associated exchangeable array. Namely: 
 %
\[f(q,x,y,u) = \ind \{u \leq w(x,y)\}.\]
 %
 \noindent Hence we hope to reinterpret the cut metric in terms of a metric on these representation functions $f$, and from there find a way to extend this metric to arbitrary exchangeable array processes. First, a simple observation.
%
\begin{align*}
    \delta_\Box(w, w') & := \sup_{S,T \subseteq [0,1]} \left | \int_{S \times T} 
 w(x,y) - w'(x,y) \; dx \; dy\right| \\
 & = \sup_{R, S,T \subseteq [0,1]} \left |\int_{R \times S \times T} 
  \int_0^1 f(q, x,y, u) - f'(q, x,y, u)   \;du \; dy \; dx \; dq\right|.
\end{align*}
%
\noindent This suggests a natural generalization to our exchangeable array processes. Given representation function $f, f'$ of d-dimensional exchangeable arrays define the following:

\[\delta(f,f') := \sup_{T_{e} \subseteq [0,1]} \bigg | \int_{\prod_e T_e} \int_0^1 \;f - f' \; \bigg|.\]

\noindent where the supremum is over  proper subsets $e \subset [d]$. \\

\daniel{So far, it looks like maybe this should be the supremum over subsets of $[0,1]^{2^d-1}$, rather than products of subsets of $[0,1]$.}

The asymmetry in the top coordinate may seem strange at first glance. We will show that, in fact, the topology is the same as the topology generated by the following symmetric metric:
\begin{align*}
\delta_s(f,f') &:= \sup_{T \subseteq [0,1]^{2^d}} \bigg | \int_T f - f' \; \bigg| \\
&= \delta_s(f,f') = \sup_{0 \leq g \leq 1} \bigg | \int_{[0,1]^{2^d}} g(f - f') \; \bigg|.
\end{align*}
where the second equality is because indicator functions are the extreme points of $\{ g \in L^2([0,1]^{2^d}) : 0 \leq g \leq 1\}$.

\daniel{Can we slacken the requirement on the set $T$ using the middle symmetry of $f$? I don't think we can in general say that it's a product of subsets of $[0,1]$, but we can probably say \emph{something}.}


\zack{This makes sense for binary arrays certainly, or if the state space is like $[0,1]$. But for our isomorphism class state spaces we would have to be a bit more careful. Sure we could label the classes as like $1,2,...$ and just integerate this, but we would want the distance to be invariant under different choices of the labeling scheme. Maybe a better idea is to use some sort of hamming metric where we instead integrate $\ind \{f \neq f'\}$. The problem with this is that it is unsigned (e.g. with graphons the integrand can be both positive or negative) so the maximizing subsets will just be all of $[0,1]$. This wouldn't quite reduce to the cut metric. Again you could try to associate some sort of signed hamming metric but it is not clear how to do so in a way that is invariant under relabelings of the states.}

\zack{Still not quite sure this is the proper metric. We originally wanted to define the supremum over every integrand, but in the case of grahpons it does not reduce to the standard cut metric in this case. I gave the counterexample of the two grahphons $w \equiv 0.5$ and $w' = \ind_{[0, \epsilon]^2}$ which is maximized on $[0,1]^3 \times [0,0.5]$}



\zack{Just like how graphons are really equivalence classes modulo Lebesgue isomorphisms, we will have to think of our representation functions $f$ in the same way, possibly where we can permute each ``part'' of the representation by a different Lebesgue isomorphism.}


\noindent Another natural class of metrics is of course just the $L^p$ norms. Namely:

\[\delta_p(f,f') := \left(\int_{\R^{2^k}} |\rho(f,f')|^p \; \right)^{1/p}.\]

\noindent where $\rho(f,f') = |f-f'|$ for $\R$-valued functions, or a hamming metric for isomorphism classes. Let's proceed with some examples.


\begin{example}[Perfectly Dependent]
In the case every element of the array is just the same random variable, the representation function $f(U, U_i,...)$ just takes the form $f(U)$ for some $f : [0,1] \to S$. Thus, the metric reduces to:

\[\delta(f,f') =  \sup_{T \subseteq [0,1]} \bigg| \int_T f(x) -  f'(x) \; dx \bigg|.\]

\zack{$\delta$ seems analogous to sort of a TV distance with respect to the measures $f \; dx, f' \; dx$ but these are not probability measures in general.}
\end{example}


\begin{example}[IID Entries]
If the array has iid entries, its representation takes the form $f(U, U_i,... , U_e) = f(U_e)$ for some $f:[0,1] \to S$. In this case, the metric reduces to:

\[\delta(f,f') =  \bigg| \int_0^1 f(x) - f'(x) \; dx \bigg|.\]

\zack{In the case of two Bernoulli random variables with success probabilities $p,q$ this just reduces to $|p-q|$}
\end{example}

\begin{example}[Polya's Urn]
In the two-color case starting with $n,m$ balls of each color we have $f(x,y) = \ind \{y \leq f(x)\}$ where $f$ is the quantile function of $Beta(n,m)$. Then the metric reduces to:

\[\delta(f,f') =  \sup_{T \subseteq [0,1]} \bigg| \int_T f(x) -  f'(x) \; dx \bigg|.\]

\noindent analogous to the perfectly dependent case.\\
\end{example}

\noindent For all three previous examples, our $L^p$ norm reduces to:

\[\delta_p(f,f') = \left(\int^1 | f(x) - f'(x)|^p \; dx \right)^{1/p}.\]

\noindent In the case $S = \R$ then $f$ is just the quantile function of the distribution. Thus, $\delta_p$ reduces to the p-Wasserstein distance.\\


\noindent Of course, the representation $f$ of a given exchangeable array need not be unique. This is even apparent in the theory of graphons, which are defined as isomorphism classes of functions $w:[0,1]^2 \to [0,1]$ modulo Lebesgue isomorphisms: $w(\cdot, \cdot) \sim w(\lambda(\cdot), \lambda(\cdot))$. The idea is that applying a Lebesgue isomorphism to a uniform does not change its distribution, and thus will not influence the overall sampling distribution. More generally, the representation functions of exchangeable arrays have a similar uniqueness result ``up to Lebesgue preserving maps''. 

The slight technicality is that we have to allow such Lebesgue preserving maps to depend on the preceding coordinates. Namely, say a map $T:[0,1]^{2^d} \to [0,1]$ \textit{preserves $\lambda$ in the last coordinate} if for fixed $x_\empty, (x_i)_{i \in [d]}, ..., (x_e)_{e \in [d]^{(d-1)}}$ the map:

\[y \mapsto T(x_\empty, (x_i)_{i \in [d]}, ..., (x_e)_{e \in [d]^{(d-1)}}, y),\]

\noindent preserves the Lebesgue measure.


For a proof of the following, see Theorem 7.28 in \cite{KallenbergProbabilisticSymmetries}, which proves a more general statement that does not restrict the array to a finite dimension $d$. For something easier to parse, the case $d=2$ can also be found in Theorem 2.1' of \cite{kallenberg_rep_thm}.
%
\begin{theorem}[Uniqueness of Representation Functions]
    The following are equivalent:
    \begin{enumerate}
        \item $f,g:[0,1]^{2^d} \to S$ generate the same exchangeable array (in distribution).

        \item There exists measurable functions $T_i,T_i' : [0,1]^{2^i} \to [0,1]$  for  $i = 0,..., d$ which preserve $\lambda$ in the last coordinate so that a.s.:
        %
        \begin{align*}
          f\Big( & T_0(x_\emptyset), (T_1(x_\emptyset, x_i))_{i \in [d]}, \dots, T_d(x_\emptyset, (x_i)_{i \in [d]}, \dots, x_{[d]}) \Big)  \\
          & = g\Big(  T'_0(x_\emptyset), (T'_1(x_\emptyset, x_i))_{i \in [d]}, \dots,  T'_d(x_\emptyset, (x_i)_{i \in [d]}, \dots, x_{[d]}) \Big).
        \end{align*}
    \end{enumerate}
    \label{thm:uniqueness_of_rep_functions}
\end{theorem}
%
An immediate corollary of Theorem \ref{thm:uniqueness_of_rep_functions} is that, in a sense, the d-uniform hypergraph chain mentioned in the previous examples is the \textit{only} binary exchangeable array process.

\begin{corollary}[Binary Representation Functions]
    If $f:[0,1]^{2^d} \to \mathbb{Z}_2$ is the representation function of a binary exchangeable array, then up to the uniqueness of Theorem \ref{thm:uniqueness_of_rep_functions}, the function $f$ is of the form:

    \[f = \ind \Big\{x_{[d]} \leq g\left(x_\emptyset, (x_i)_{i \in [d]}, ..., (x_e)_{e \in [d]^{(d-1)}} \right)\Big\}.\]

    \label{cor:classification_binary_array_processes}
\end{corollary}

Again, we are primarily interested in the extremal representation functions, which by Theorem \ref{thm:boundary_is_ergodic_arrays}, are the ones depending trivially on $x_\emptyset$. Hence in the $d=2$ case, for extremal representation functions we exactly recover the uniqueness statement for graphons discussed above.
 
 Two arrays generated by equivalent representation functions have the same distribution, and any worthwhile metric on such representation functions ought to assign distance zero between such functions. This suggests if all we care about is the sampling distribution then, just as is done with graphons, we ought to equate all functions whose arrays have the same law and instead work with equivalence classes of representation functions. We formalize this with the following:

\[d(f, f') := \inf_{g \sim f, g' \sim f'} \delta(g,g') \quad \red{= \inf_{g \sim f} \delta(g,f')}.\]

\noindent As with the theory of graphons, the notion of convergence induced by this metric is the same as that of the embedding densities as well as the Doob-Martin topology.

\begin{theorem}[Convergence via the Cut Metric]
    For representation functions $g_n,g$, the following are equivalent.
    \begin{enumerate}[label = (\roman*)]
        \item $emb^W(F,g_n) \to emb^W(F,g)$ as $n \to \infty$ for all finite arrays $F$.

        \item $\delta(g_n,g) \to 0$ as $n \to \infty$.
        
        \item $\delta_s(g_n,g) \to 0$ as $n \to \infty$.

    \end{enumerate}
    \label{thm:conv_embed_implies_conv_cut_metric}
\end{theorem}

\daniel{I'll write down my attempt at proving the equivalence here:}

First, we show that the embedding densities are Lipschitz continuous with respect to the cut metric, and thus that convergence in the cut metric implies convergence in the embedding densities. This shows the cut topology is finer than that of the embedding densities. Many of these proofs are direct extensions of those in \cite{Meliot2024} and \cite{borgs2007convergent}.

\begin{lemma}[Cut Metric Controls Embedding Densities] \label{lem:metric_stronger_than_embeddings}
    Let $F$ be a finite, binary $d$-array of size $k \geq d$, and let $f,f' : [0,1]^{2^d} \to \{0,1\}$ be middle symmetric functions. Then
    $$|emb^W(F,f) - emb^W(F,f')| \leq |\mathbf 1(F)| \delta(f,f') \leq |\mathbf 1(F)| \delta_s(f,f').$$
\end{lemma}

\noindent The next step is to show that convergence in the embedding densities implies convergence in the cut metric . This follows the same proof strategy as the analogous theorem for graphons. Namely, there are a few layers of approximations:

\begin{enumerate}
    \item Approximate the representation function $f$ by a large finite (random) array $X_n(f)$ sampled from it.

    \item Show when the embedding densities of $f,g$ are close, the sampled arrays are close as well.

    \item Approximate a representation function $f$ by a ``step function'' with few parts.
\end{enumerate}

\noindent The first step requires us to uniquely associate a specific representation function $f_x$ to every finite array $x$. This allows us to embed our finite arrays in our space of representation functions. A natural candidate is just a continuous version of the hypergraph adjacency matrix, analogous to how we associate a graphon to a finite graph:

\[f_x(u_\empty, (u_i)_{i \in [d]}, ..., u_{[d]}) := \ind \left\{(\lceil |x| \cdot u_1 \rceil,..., \lceil |x| \cdot u_d \rceil) \in edges(x) \right\}.\]

\noindent We call $f_x$ the \textit{sampling function} associated to a finite array $x$. For this embedding to respect the topology of our embedding densities, we require $emb^W(F, f_x) = emb^W(F, x)$. Namely, the value of the (weak) embedding density does not change whether we view $x$ as an array or as its corresponding representation function. It is easy to verify that $f_x$ satisfies this constraint:
%
\begin{align*}
    emb^W(F, f_x) & := \int_{[0,1]^{2^k}} \prod_{S \in 1(F)} f_x(y_{(S)}) \; d\lambda_{2^k}(y) \\
    %
    & := \int_{[0,1]^{d}} \prod_{S \in 1(F)} \ind \left\{(\lceil |x| \cdot y_{s_1} \rceil,..., \lceil |x| \cdot y_{s_d} \rceil) \in edges(x) \right\} \; d\lambda_{d}(y) \\
    %
    & := \int_{[0,1]^{d}}  \ind \left\{ (i \mapsto \lceil |x| \cdot y_i \rceil)  \in emb(F,x)  \right\} \; d\lambda_{d}(y) \\[5pt]
    %
    & = emb^W(F,x).
\end{align*}
%
This gives us the desired embedding. We will come back to it later. 

The following lemma and corollary handle step two above by relating the distribution of our sampled arrays to the embedding densities.


\begin{lemma}[Sampling Probability Formula] \label{lem:automorphism_formula_for_sampling_probabilities}
    Let $A_n$ be a finite, binary array of size $n$ sampled from $f$. Then 
    $$\mathbb P(A_n = x) = \frac{n!}{|\operatorname{Aut}(x)|^2} \sum_{x' \geq x} (-1)^{|1_x| - |1_{x'}|} emb^W(x',f) \red{ \; = \frac{n!}{|\operatorname{Aut}(x)|^2} \cdot emb(x,f)}.$$
\end{lemma}

\noindent The next result shows that if the weak embedding densities of two representation functions are close, the finite sampling distributions are close as well. 

\begin{corollary}[Total Variation Bound for Samples] \label{cor:tv_bound_between_samples}
    Let $A_n,A_n'$ be finite, binary $d$-arrays of size $n$ sampled from $f,f'$, respectively. Then
    $$\| \operatorname{Law}(A_n) - \operatorname{Law}(A_n') \|_{\operatorname{TV}} \leq \frac{n!}{2} \sum_{|x| = n} \sum_{x' \geq x} |emb^W(x',f) - emb^W(x',f')|,$$
    where this sum is over pairs of arrays of size $n$.
\end{corollary}

\begin{lemma}[Coupling of Samples] \label{lem:coupling_of_samples}
    Let $(A_n)_n,(A_n')_n$ be sequences of finite arrays sampled from $f,f'$, and let $f_n = f_{A_n},f_n' = f_{A_n'}$ be the associated sampling functions. Then there exists a coupling of $A_n,A_n'$ such that
    $$\mathbb E[\delta_s(f_n,f_n')] \leq \frac{n!}{2} 3^{\binom{n}{d}} \max_{|x| = n} |emb^W(x,f) - emb^W(x,f')|.$$
\end{lemma}

This completes the approximation specified in part two above. We now state a technical lemma which is at the heart of parts one and three, postponing its proof until later.

\begin{lemma}[Approximation by Sampled Arrays] \label{lem:finite_functions_converge_to_overall_sampling_functions}
    Let $(A_n)_n$ be a sequence of finite arrays \red{of size $n$} sampled from $f$, and let $f_n = f_{A_n}$ be the associated sampling functions. Then
    $$\mathbb E[\delta_s(f_n,f)] \leq \frac{5}{\sqrt{\log_2 n}}.$$
\end{lemma}

\daniel{This proof will require some more work, like a version of the graph regularity lemma.}

\noindent Now we are ready to prove the main result. \noindent Note that no part of the proof above utilizes the specific definition of the cut norm or the sampling function associated to a finite array. All that is left now is to prove Lemma \ref{lem:finite_functions_converge_to_overall_sampling_functions}.

\daniel{Here, I'll collect the ingredients and arguments needed to prove Lemma \ref{lem:finite_functions_converge_to_overall_sampling_functions}. We can decide later how we want to organize all these results.}

Given a partition $\mathcal P$ of $[0,1]$, define the grid $\sigma$-algebra on $[0,1]^d$ by $\mathcal G_{\mathcal P} := \sigma(\mathcal P)^{\otimes d}$. Martingale convergence tells us that $\E[f \mid \mathcal G_{\mathcal P}] \to f$ as $\mathcal P$ becomes increasingly refined. We will, however, need a quantitative rate of convergence in terms of the number of pieces in the grid.

\begin{lemma}[Step-Function Approximation] \label{lem:approximation_by_step_fn}
    For any $\varepsilon > 0$, there exists a measurable partition $\mathcal P$ with $q \leq [placeholder]$ pieces such that
	    $$\delta_s(\E[f \mid \mathcal G_{\mathcal P}],f) \leq \varepsilon \| f \|_{L^2([0,1]^{2^d})}.$$
\end{lemma}

\daniel{For the following proof, we need to have defined the metric as being the inf over all relabelings (i.e. Lebesgue isomorphisms), since  the strategy is to find a relabeled version of $g_n$ that matches very closely with $g$. In the proof, I'll just assume that we've defined $\delta_s,\delta$ that way. We can sort out the notation later.}

\zack{So it seems like there are two problems. The first is clearly proving Lemma \ref{lem:approximation_by_step_fn} which says we can approximate our representation function $f$ by a step function (which uses only the coordinates associated to the singletons). The second is proving that this step function actually well-approximates $f$ in the cut norm. It seems the part showing the step functon}


% Cut metric for Nonbinary Arrays %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 
\subsection{Cut metric for nonbinary arrays}

\zack{This section is very rough for now. }

First, we need to extend the notion of a weak embedding to the nonbinary case. For a subset of states $C \subseteq S$ define a \textit{C-weak embedding} of $x$ into $y$ to be an injective map $\phi: [|x|] \to [|y|]$ which preserves the states in $C$. Namely, if $s \in C$ then $x_e = s$ implies $y_{\phi(e)} = s$. In the case $C = \{1\}$ and $S = \{0,1\}$ this reduces to the standard notion of weak embeddings of a binary array. Denote by $emb_C(x,y)$ the set of all such C-weak embeddings and $emb^C(x,y)$ the corresponding embedding density. Then we have the following extension of Lemmas \ref{lem:relating_different_notions_of_embedding} and \ref{lem:equivalence_of_dm_and_density_convergence}. 

\zack{The above gives us a way of considering multiple states to be the same. Namely, if we clump all the states in $C^c$ together it does not change the $C-weak$ embedding density.}


\begin{lemma}[Weak Embeddings for Nonbinary Arrays]
    For any $C \subseteq S$ then:

    \begin{enumerate}[label=(\roman*)]
        \item $emb^C(x,y) = \sum_{x' \geq_C x} emb(x', y)$

        \item $emb(x,y) \leq \sum_{x' \geq_C x} \red{todo} $
    \end{enumerate}

    \noindent Moreover, if $C = S \setminus \{s\}$ then:

    \begin{enumerate}
        \item[(iii)]  $emb(x,y) = \sum_{x' \geq_C x} (-1)^{|1_s(x)| - |1_s(x')|} \cdot emb^C(x', y)$
    \end{enumerate}
    

    \noindent where $x' \geq_C x$ means $x'$ can be obtained from $x$ by changing some of the values $x_e$ with $x_e \notin C$ and $1_s(x) = | \{e \; : \; x_e = s\} |$ is the number of entries of $x$ which are $s$. 
    
    In particular, this shows that convergence of $emb(\cdot, y_N)$ implies convergence of $emb^C(\cdot, y_N)$ for any $C$, and the converse is true if $C = S \setminus \{s\}$. 

    \label{lem:weak_embeddings_for_nonbinary_arrays}
\end{lemma}

\zack{Note no result (iii) can hold for smaller $C$ because then $x' \geq_C x$ fails to be a partial order, so we cannot find the standard Mobius transform.}



\noindent There is a simple way of encoding a nonbinary array with finite state space $S$ as a collection of binary arrays of the same dimension: simply apply a one-hot encoding to the state space. More specifically, for each $A \subseteq S$ define the function $g_A: S \to \{0,1\}$ by $g_A(s) = \ind \{s \in A\}$. Thus each $d-$dimensional array with values in $S$ is encoded as $2^{|S|}$ binary arrays of the same size. First, we show there is a relationship between embeddings in the original state space and embeddings of the corresponding binary arrays.

\zack{This might be overkill. Maybe just sets of the form $A = \{s\}$ suffice.}


% Lemma %
\begin{lemma}[Binary Projections of Nonbinary Arrays]
    For d-dimensional arrays $x,y$ with values in some finite set $S$ then:
    %
        \[\text{ind}(x,y) = \bigcap_{s \in S} \text{ind}(g_s(x), g_s(y)).\]
    %
    \noindent \red{Moreover, $emb( \cdot ,y_N) \to emb( \cdot , f)$ as $N \to \infty$ if and only if $emb(\cdot, g_A(y_N)) \to emb(\cdot, g_A \circ f)$ for all $A \subseteq S$.} 

    \label{lem:reducing_nonbinary_conv_to_binary}
\end{lemma}

Given an exchangeable array process $(X_e)$ represented by $f$, lemma \ref{lem:state_equivalence} implies the array process $(g_A(X_e))$ is a binary exchangeable array process represented by $g_A \circ f$. Thus it is natural to define:
%
\begin{equation}
    \delta(f,f') := \frac{1}{2^{|S|}} \sum_{A \subseteq S} \delta(g_A \circ f, \; g_A \circ f').
\end{equation}
%
\noindent whenever $f,f'$ share the same state space $S$. 


\zack{Can we define it when $f,f'$ have different state spaces? Maybe just extend the smaller state space to the larger one by adding states it never takes on. }

\daniel{You an do that, but I think the distance will always be 1. But if you can make sense out of there being some states which are considered the same between the two combinatorial structures, that might be interesting.}

\daniel{One thing that might be a good idea, as well, is to cover the case of structures with interactions of different orders, thinking of a product of arrays of dimension 1, 2,...,d (with a product space metric). Maybe that's not necessary, since there might be a way to encode that in the state space by enlarging $S$, but if we did a product of arrays of dimension $d$ for all $d \in \N$, we could represent hypergraphs that way. The metric could probably be taken to be something like $\sum_{d=1}^\infty 2^{-d} \text{dist}_d$.}


% Doob-Martin Background Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Background on Doob-Martin Boundary Theory}
\label{sec:doob_martin_background_appendix}
Here we aim to provide some of the essentials of Doob-Martin boundary theory. We restrict our attention to the case of discrete Markov chains relevant to this paper. The primary reference on this topic is \citep{Doob_original} but \citep{sawyerMB} and \citep{DM_Remy} also provide good overviews for the discrete case in more modern terminology.\\
%
\indent Suppose $\Xn$ is a transient Markov chain with a countable state space $E$ and transition kernel $p$. Define the \textit{Green's kernel} or \textit{potential kernel} $g: E \times E \to \mathbb{R}_{\geq0}$ by:
%
\begin{equation}
    g(x,y) := \sum_{n \geq 0} p^n(x,y) = \E_x (\text{number of times $X_n$ hits y}).
    \label{eqn:green's_kernel}
\end{equation}
%
\noindent Assume further there exists a base state $x_0$ from which the chain is irreducible, namely $g(x_0,y) > 0$ for all $y \in S$. For this paper, $x_0$ will typically be the empty array. Note that transience implies that $g(x,y) < \infty$. In this paper, we will typically be interested in \textit{graded} chains in which the state space $E  =\bigcup E_n$ can be partitioned in such a way so that transitions exclusively occur from $E_n$ to $E_{n+1}$. In this case, we can think of $g(x,y)$ as $\P_x(X_n \text{ hits }y)$. Using this, define the \textit{Doob-Martin kernel} $K:E \times E \to \R_{\geq 0}$ with respect to base state $x_0$ to be:
%
\begin{equation}
    K(x,y) := \frac{g(x,y)}{g(x_0,y)} = \frac{\P_x(X_n \text{ hits y})}{\P_{x_0}(X_n \text{ hits y})}. 
\end{equation}

%
\noindent Our irreducibility assumption implies this is well-defined. Observe that for each fixed $y$ the function $K(\cdot, y)$ is ``almost harmonic'' in the sense that:

\begin{equation*}
    \sum_{z \in S} p(x,z)K(z, y) = K(x,y) \quad x \neq y. 
\end{equation*}


\noindent Intuition might suggest that if we can let $y$ ``go off to infinity'' we ought to get a truly harmonic function, and this suggests an investigation of sequences of states $\{y_n\}$ in $S$ so that the sequence of real numbers $\{K(x, y_n)\}$ converges for all $x\in S$. The Doob-Martin boundary will make this intuition precise. 

\indent Note that for each fixed $x$ that $K(x, y)$ is a bounded function of $y$. To see why, note irreducibility implies $p^m(x_0, x) > 0$ for some $m$ so that: 
%
\[0 \leq K(x,y) := \frac{g(x,y)}{g(x_0, y)} \leq \frac{g(x,y)}{p^m(x_0, x) g(x,y)} = \frac{1}{p^m(x_0,x)} =: C_x.\]

%
\noindent Hence we can view the collection of functions $\{K(\cdot, y) \}_{y \in E}$ as a subset of $\bigtimes_{x \in E} [0, C_x] \subseteq \R^E_{\geq0}$. As $\bigtimes_x [0,C_x]$ is compact under its standard product topology, then $\{K(\cdot, y) \}_{y \in E}$ is precompact. Moreover, as $K(\cdot, y) \neq K(\cdot, y')$ for $y \neq y'$ we can identify $E$ with $\{K(\cdot, y) \}_{y \in E}$. Its closure $\overline{E} \subseteq \R^E_{\geq 0}$ under this identification is thus compact. We call $\overline{E}$ the \textit{Doob-Martin compactification} of $E$ and its boundary $\partial E := \overline{E} - E$ the \textit{Doob-Martin boundary}. In summary, we have turned $E$ into a metric space by identifying it with a subset of $\R^{E}_{\geq 0}$ which already carries a natural metric structure, and used this identification to compactify $E$. In fact, this can be done with any class of functions $\cF$ that separates points. Here we just choose $\cF = \{K(\cdot, y)\}_{y \in E}$.

%
\begin{remark}
    By construction, a sequence $\{y_n\} \in E$ converges to a point $\alpha \in \overline{E}$ if and only if $\{K(\cdot, y_n)\}$ converges in $\R^E_{\geq 0}$, namely if and only if $K(x, y_n)$ converges in $\R_{\geq 0}$ for all $x \in E$. Moreover, for each fixed $x$ the function $K(x, \cdot) : E \to \R_{\geq 0}$ extends continuously to $\overline{E}$. 
    \label{rmk:conv_in_dm}
\end{remark}
%
%
\noindent There is a close relationship between the Doob-Martin kernel $K$ and the backwards transition probabilities of the chain. Namely, for $x \in E_n, y_N \in E_{N}$ by applying Bayes rule:
%
\[K(x,y) = \frac{\P(X_{N} = y_N \; | \; X_n = x)}{\P(X_{N} = y_N)} = \frac{\P(X_n = x \; | \; X_{N} = y_N)}{\P(X_n = x)}.\]
%
\noindent Thus $K(x, y_N)$ converges as $N \to \infty$ for all $x$ if and only if the conditional probabilities $\P(X_n = x \; | \; X_N = y_N)$ converge for all $x$. 


%%%%%%%%%%%%%%%%%% Minimal Boundary and representation of harmonics %%%%%%%%%%%%%%
\subsection{Minimal Boundary and Representation of Harmonics}

\noindent Essentially by construction we have $K(\cdot, \alpha)$ is harmonic for every $\alpha \in \partial E$. 
\noindent A specific subset  $\partial_m E \subseteq \partial E$, the \textit{minimal boundary}, will play a crucial role in our understanding of harmonic functions and the Doob-Martin boundary. It is defined as:

\begin{equation}
    \partial_m E := \{ \alpha \in \partial E : K(\cdot, \alpha) \text{ is minimal harmonic}\}.
\end{equation}

\noindent Here, a \textit{minimal harmonic} function is a harmonic function $u \geq 0$ so that for any other harmonic $w$ satisfying $0 \leq w \leq u$ we must have $w = c u$ for some constant $c \geq 0$.
%
Let $\Hn$ denote the set of nonnegative harmonics with $h(x_0) = 1$. Note that $\Hn$ is a compact, convex set whose extreme points are precisely the minimal harmonic functions, and again clearly $K(\cdot, \alpha) \in \Hn$. Thus the minimal boundary is precisely just the subset of $\alpha$ so that $K(\cdot, \alpha)$ is an extreme point of $\Hn$. Ideas from convex analysis suggest we can represent other harmonic functions as an integral over these extreme points, and the following representation theorem shows this is indeed true. For every $u \geq 0$ harmonic there exists a unique measure $\mu_u$ supported on $\partial_m E$ satisfying:

\begin{equation}
    u(x) = \int_{\partial_m E} K(x, \alpha) \mu_u(d\alpha).
    \label{eqn:harmonic_representation_theorem}
\end{equation}

\noindent Moreover, as $K(\cdot, \alpha)$ is a harmonic for $\alpha \in \partial_m E$ and $|K(x, \alpha)| \leq C_x$ then equation \eqref{eqn:harmonic_representation_theorem} defines a nonnegative harmonic function for any finite measure $\mu_u$ which satisfies $u(x_0) = 1$ if and only if $\mu_u$ is a probability measure. In words, as $K(\cdot, \alpha)$ can be thought of as points on the minimal boundary, nonnegative harmonic functions are characterized by their distribution on the boundary. For the special case of $u = K(\cdot, \alpha)$ for $\alpha \in \partial_m E$ clearly we have $\mu_{u} = \delta_\alpha$. A full proof of these facts can be found in \cite{sawyerMB}.\\

\noindent In some cases such as the R\'emy chain studied in \cite{DM_Remy}, $\partial E = \partial_m E$, but in general this is not true for an arbitrary Markov chain.


%%%%%%%% h processes 
\subsection{Using h-processes to Condition at Infinity} \label{sec:h_processes}

 Given a harmonic function $h \in \Hn$, irreducibility and harmonicity imply in fact that $h > 0$ and so we can define a new transition mechanism $p^h$ which ``tilts'' the dynamics using $h$:
%
\begin{equation}
    p^h(x,y) := \frac{1}{h(x)} p(x,y) h(y), \quad x,y \in E.
    \label{eqn:h-process}
\end{equation}
%
\noindent In particular, $p^h(x_0, y) = p(x_0,y)h(y)$. Thus the new chain reweights transitions of the original chain by a factor of $h(y) / h(x)$, favoring transitions towards states where $h$ is large, and in the case $h \equiv 1$ we simply obtain the original transition mechanism. We call the new Markov chain $X^{(h)}_n$ with transition probabilities $p^h$ the \textit{Doob h-process} or \textit{Doob h-transform} of $X_n$ corresponding to $h(x)$.
The $h$-processes $X_n^{(K(\cdot, \alpha))}$ are of natural interest because they characterize the distributions of all other $h$-processes, including the original chain:
%
\begin{equation}
\P_{x_0}^h( B) = \int_{\partial_m E} \P_{x_0}^{K(\cdot, \alpha)}(B) \; \mu_h(d\alpha). 
    \label{eqn:K(., alpha) processes characterize h processes}
\end{equation}%
%
\noindent for some unique measure $\mu_h$ supported on the minimal boundary. In particular, these $h$-processes converge a.s. in the Doob-Martin topology to the minimal boundary and have limiting distributions:
%
\begin{equation}
    \P_{x_0}(\lim X_n^{(h)} \in A) = \int_{A \cap \partial_m E} K(x, \alpha) \mu_h(d\alpha),
    \label{eqn:limit_distribution_h-process}
\end{equation}
\begin{equation}
    \P_{x_0}(\lim X_n \in A) = \int_{A \cap \partial_m E} K(x, \alpha) \mu_1(d\alpha). 
    \label{eqn:limit_distribution_original_chain}
\end{equation}
%
\noindent Here $\mu_1$ is the measure corresponding to $h \equiv 1$. Namely, in the setting of Doob-Martin boundaries an $h$-process is merely a chain conditioned to have distribution $\mu_h$ on the minimal boundary. Again, if $\alpha \in \partial_m E$ then $\mu_{K(\cdot, \alpha)} = \delta_\alpha$ so $\lim X_n^{K(\cdot, \alpha)} = \alpha$ almost surely. Thus we can truly think of the chain $X^{(K(\cdot, \alpha))}_n$ to be our original Markov chain conditioned to equal $\alpha \in \partial_m E$  ``at infinity''. Note we can extend the above to include measures on the full boundary, but then the measures $\mu_h$ are no longer necessarily unique. 


% Bridges and its relationship to the Doob martin boundary 
\subsection{Infinite Bridges} \label{sec:infinite_bridges}

 \indent A finite-step Markov process $(Y_n)_{n = 1}^N$ with $Y_0 = x_0$ which has the same backwards transition probabilities as $(X_n)_{n \in \N}$ is called a (finite) \textit{bridge} for $X_n$. Similarly, if the index set is $\N$ we call $(Y_n)_{n \in \N}$ an \textit{infinite bridge}. One key example of a bridge will be the following conditioned chain. Let $(X^{(y)}_0, X^{(y)}_1, ..., X^{(y)}_{N(y)})$ be the finite-step chain obtained from $X_n$ by conditioning on $X_{N(y)} = y$. Then we see:
%
 \begin{align*}
     \P(X^{(y)}_n = i | X^{(y)}_{n+1} = j) & =  \frac{\P_{x_0}(X_n = i, X_{n+1} = j, X_{N(y)} = y)}{\P_{x_0}( X_{n+1} = j, X_{N(y)} = y)} \\
     %
     & =  \frac{\P_{x_0}(X_n = i, X_{n+1} = j) \cdot \P_{j}( X_{N(y)} = y)}{\P_{x_0}( X_{n+1} = j) \cdot \P_j( X_{N(y)} = y)} \\
     %
     & =  \P_{x_0}(X_n = i | X_{n+1} = j).
 \end{align*}
% 
 Namely, $X^{(y)}_n$ is a finite bridge of $X_n$. Moreover, 
% 
 \begin{align*}
     \P(X^{(y)}_{n+1} = i | X^{(y)}_n = j) & =  \frac{\P_{x_0}(X_n = j, X_{n+1} = i, X_{N(y)} = y)}{\P_{x_0}( X_n=j, X_{N(y)} = y)} \\
     % 
     & = \frac{\P_{x_0}(X \text{ hits j}) \cdot p(j,i) \cdot \P_i(X \text{ hits y})}{\P_{x_0}(X \text{ hits j}) \cdot \P_j(X \text{ hits y})} \\ 
     %
     & = K(j,y)^{-1}\cdot p(j,i) \cdot K( i, y).
 \end{align*}

\noindent Observe that this is essentially the definition of an $h$-process with $h = K(\cdot, y)$, except for the fact that $K(\cdot, y)$ is not harmonic for $y \in S$. However, as before, intuition suggests we ought to take the limit as $y$ ``goes to infinity'' to get a truly harmonic function $K(\cdot, \alpha)$. Namely, by Remark \ref{rmk:conv_in_dm} we know a sequence $(y_k)$ in $S$ converges to a point $\alpha \in \partial S$ if and only if $K(x, y_k)$ converges in $\R_{\geq 0}$ to $K(x, \alpha)$ for all $x \in S$ which of course implies:

\[\P(X^{(y_k)}_{n+1} = i | X^{(y_k)}_n = j) \xrightarrow{k\to \infty} K(j,\alpha)^{-1}\cdot p(j,i) \cdot K( i, \alpha) =: p^{K(\cdot, \alpha)}(j,i).\]


\noindent As observed in \cite{hans_bridges} such a sequence converges in the Doob Martin topology if and only if finite initial segments of the corresponding bridges converge in distribution, and two such sequences converge to same limits if and only if their finite initial segments converge to same limit as well. Namely, convergence of the corresponding bridges characterizes convergence in the Doob-Martin boundary, so if we can understand these bridges we can understand the Doob-Martin boundary. To be precise, by convergence of finite initial segments we mean for each finite $M$ we have: \\
%
\begin{equation}
    (X^{(y_k)}_0, X^{(y_k)}_1, ..., X^{(y_k)}_M) \Rightarrow (X^{K(\cdot, \alpha)}_0, X^{K(\cdot, \alpha)}_1, ..., X^{K(\cdot, \alpha)}_M)
    \label{eqn:2.3_initial_sections_finite_bridges}
\end{equation}
%
\noindent as $k \to \infty$. The Doob h-transform provides a bijective correspondence between nonnegative harmonic functions with $h(x_0) = 1$ and infinite bridges. Namely, given such a harmonic $h$, the Doob h-transform of the original chain is certainly an infinite bridge as:
%
\begin{align*}
  \P(X^h_n = x \; | \; X^h_{n+1} = y) & = \frac{\P(X^h_{n+1} = y \; | \; X^h_{n} = x) \cdot \P(X^h_n = x)}{\P(X^h_{n+1} = y)} \\
  %
  & = \frac{\frac{h(y)}{h(x)}\P(X_{n+1} = y \; | \; X_{n} = x) \cdot \frac{h(x)}{h(x_0)}\P(X_n = x)}{\frac{h(y)}{h(x_0)}\P(X^h_{n+1} = y)} \\
  %
  & = \P(X_n = x \; | \; X_{n+1} = y).
\end{align*}



\noindent Likewise, given a bridge $X^\infty_n$ then $\eqref{eqn:K(., alpha) processes characterize h processes}$ tells us there exists a measure $\mu_h$ so that $X^\infty_n$ is a mixture of the extremal bridges $\P^{K(\cdot, \alpha)}$. Then the harmonic $h$ corresponding to $X^\infty_n$ is just a corresponding mixture of these extremal harmonics $K(\cdot, \alpha)$:
%
\[h(x) = \int_{\partial_m E} K(x, \alpha) \; \mu_h(d \alpha).\]
%
\indent Call an infinite bridge \textit{extremal} if it has a trivial tail $\sigma$-field, or equivalently if it cannot be written as a nondegenerate mixture of other infinite bridges. By \eqref{eqn:K(., alpha) processes characterize h processes} and the correspondence between bridges and harmonic functions, all infinite bridges are just mixtures of the extremal bridges. Hence characterizing the extremal bridges allows us to characterize all infinite bridges.

Our observations from Section \ref{sec:h_processes} tell us that for $\alpha \in \partial_m S$ that the conditioned chain $X_n^{K(\cdot, \alpha)}$ is extremal as it converges almost surely to the constant $\alpha$. Moreover, the converse is true: $X^{K(\cdot, \alpha)}$ is extremal only if $\alpha \in \partial_m S$. Hence characterizing the extremal bridges will allow us to characterize the minimal boundary, and thus by \eqref{eqn:harmonic_representation_theorem} characterize all nonnegative harmonic functions for this chain. \red{But again unless $\partial S = \partial S_m$ we won't be able to characterize the full boundary this way. We will have to allow mixtures of extremal infinite bridges}



% Proof Appendix %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Proof Appendix}
\label{sec:proof_appendix}

\begin{proof}[Proof of Lemma~\ref{lem:state_equivalence} (State-Space Projections of Exchangeable Arrays)]
The result follows immediately from applying $g$ coordinatewise. If $(X_e)$ is exchangeable, then for each finite permutation $\sigma$,
\[
(g(X_{\sigma(e)}))_{e \in \N^{(d)}} \overset{\mathrm{law}}{=} (g(X_e))_{e \in \N^{(d)}}.
\]
Moreover, if $X_e = f(U_\emptyset,(U_i)_{i \in e},\ldots,U_e)$ in distribution, then $g(X_e)$ is represented by $g \circ f$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:index_equivalence} (Markov Property of the Unlabeled Chain)]
Firstly note that the definition of an exchangeable array implies:
%
\[\P(L_n = x) = \P(L_n = x') \qquad \forall x \sim x'.\]

\noindent Moreover if $x \sim x'$ are related by $\sigma \in Sym(\N)$ then again exchangeability implies:
%
\begin{align*}
    \P(M_{n+1} = [y]_\sim,\; L_n = x) & = \sum_{y' \sim y} \P(L_{n+1} = y', \; L_n = x)\\
    & = \sum_{y' \sim y} \P(L_{n+1} = \sigma(y'), \; L_n = \sigma(x)) \\ 
    & = \sum_{z \sim y} \P(L_{n+1} = z, \; L_n = x') \\
    & = \P(M_{n+1} = [y]_\sim,\; L_n = x').
\end{align*}
%
\noindent Thus we see:
%
\[\P(M_{n+1} = [y]_\sim \; | \; L_n = x) = \P(M_{n+1} = [y]_\sim \; | \; L_n = x') \qquad \forall x \sim x', \quad [y]_\sim \in [E_{n+1}]_\sim.\]
%
\noindent Thus the transition probability from $x$ into the class $[y]_\sim$
depends only on the equivalence class $[x]_\sim$. The projected transition
kernel is therefore well-defined on equivalence classes, and hence $M_n$ is a
Markov chain.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:boundary_labeled_chain} (Boundary of the Labeled Chain)]
Clearly any deterministic array cannot be written as a mixture of other arrays and hence such arrays are extremal. Moreover, any nondeterministic array is a mixture (over its own distribution) of determinsitic arrays and hence is not extremal.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:harmonic_functions_labeled} (Exchangeable Bridges of the Labeled Chain)]
Suppose the harmonic $h \geq 0$ with $h(x_0) = 1$ is the one corresponding to this particular bridge. If $L_n$ is the original labeled exchangeable array process and $L^h_n$ is the infinite bridge:
%
\[\P(L^h_0 = x_0, ..., L^h_n = x_n) = h(x_n) \cdot \P(L_0 = x_0, ..., L_n = x_n).\]
%
\noindent The right hand side is zero if the states $x_0, ... , x_n$ are not nested in the sense of arrays, $x_m = (x_n) |_{[m]^{(d)}}$ for $m < n$, and thus the same is true of the left hand side. Thus, by taking a projective limit, the chain $L_n^h$ a.s. defines an array ($X^h_e)$ of the same dimension as the original exchangeable array $(X_e)$ of which it is the corresponding array process. 

Now if $\sigma \in Sym(\N)$ let $\tx := \sigma(x)$ be the array generated by permuting the coordinates of $x$. Then the definition of the Doob h-transform and the exchangeability of the original array imply:
%
\begin{align*}
    \P(L^h_0 = \tx_0, ..., L^h_n = \tx_n)  & = h(\tx_n) \cdot \P(L_0 = \tx_0, ..., L_n = \tx_n) \\
    %
    & = h(\tx_n) \cdot \P(L_0 = x_0, ..., L_n = x_n) \\
    & = \frac{h(\tx_n)}{h(x_n)} \cdot \P(L^h_0 = x_0, ..., L^h_n = x_n). 
\end{align*}

\noindent These two are equal if and only if $h(\tx_n) = h(x_n)$. Since $x_n$ and $\sigma$ were arbitrary, the result follows from the fact that the finite dimensional distributions of our array $(X^h_e)$ are determined from the distribution of $(L_n^h)$.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:unlabeled_associated_to_exchangeable_array} (Bridges of the Unlabeled Chain)]
The bridges of $(M_n)$ are in correspondence with harmonic functions
$\tilde{h}$ of this chain. Define $h(x) := \tilde{h}([x])$ on the labeled state
space. Since the transition kernel of $M_n$ is obtained by summing the
transition probabilities of $L_n$ over equivalence classes, harmonicity of
$\tilde{h}$ gives
%
\[
    h(x)
    =
    \tilde{h}([x])
    =
    \sum_{[y]_\sim} p_M([x]_\sim,[y]_\sim)\tilde{h}([y]_\sim)
    =
    \sum_y p_L(x,y) h(y).
\]
%
Thus $h$ is a harmonic function for the labeled chain and is constant on the
equivalence classes of $\sim$. Applying Lemma~\ref{lem:harmonic_functions_labeled}
gives the corollary.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:boundary_is_ergodic_arrays} (Extremal Bridges and Ergodic Arrays)]
The last three statements are equivalent by Lemma 7.35 of \cite{KallenbergProbabilisticSymmetries}. For the implication $(2) \to  (1)$, suppose for contraposition that we have a nonextremal bridge $M_n^{\tilde{h}}$ associated to the harmonic $\tilde{h}$. Then equation \eqref{eqn:K(., alpha) processes characterize h processes} applied to the Doob-Martin kernel $\tilde{K}$ of our unlabeled chain implies:

\[\P((M_n^{\tilde{h}}) = [x]_\sim) = \int_{\partial_m \tilde{E}} \P^{\tilde{K}(\cdot, \alpha)}([x]_\sim) \; \mu_{\tilde{h}}(d\alpha),\]

\noindent for some nondegenerate measure $\mu_{\tilde{h}}$ on the minimal boundary $\partial_m \tilde{E}$ of the unlabeled chain. By construction of the Doob-Martin boundary, the functions $\tilde{K}(\cdot, \alpha)$ are harmonic for the unlabeled chain. Thus by Corollary \ref{cor:unlabeled_associated_to_exchangeable_array} the distributions $\P^{\tilde{K}(\cdot, \alpha)}$ correspond to the unlabeled chains associated to exchangeable arrays $(X^\alpha_e)$. Likewise, $M_n^{\tilde{h}}$ corresponds to some exchangeable array $(X_e^{\tilde{h}})$ as well. Thus we can rephrase the previous equation in terms of the associated exchangeable arrays:

\[\P((X^{\tilde{h}}_e) \in [x]_\sim) = \int_{\partial_m \tilde{E}} \P(X_e^\alpha \in [x]_\sim) \; \mu_{\tilde{h}}(d\alpha).\]

\noindent These probabilities are constant on the equivalence classes by exchangeability. Hence:

\[\P((X^{\tilde{h}}_e) = x) = \int_{\partial_m \tilde{E}} \P(X_e^\alpha = x) \; \mu_{\tilde{h}}(d\alpha).\]

\noindent Thus as $\mu_{\tilde{h}}$ is nondegenerate, $(X_e^{\tilde{h}})$ is a nondegenerate mixture of other exchangeable arrays, and hence not ergodic. 

% proof of converse
For the implication $(1) \to (2)$ suppose that we have an extremal bridge $M_n^{\alpha}$ associated to the minimal harmonic $\tK(\cdot, \alpha)$, where $\alpha$ is in the minimal boundary  $\tE$. As mentioned in Section 3, this is equivalent to the fact that this bridge has a trivial tail $\sigma$-field almost surely. By Corollary \ref{cor:unlabeled_associated_to_exchangeable_array} there is an exchangeable array $(X_e^\alpha)$ associated to this bridge. Suppose this array has distribution $\nu$ and that $I$ is a measurable $\nu$-invariant event. Let $r_n$ denote the function from infinite $d$-dimensional arrays to size $n$ arrays that just restricts the entries of $(X^\alpha_e)$ to $e \in [n]^{(d)}$. Then for every finite choice of $N \in \N$ because the labeled chain $L_n^\alpha$ is just the size n ``top-left corner'' of the array $X^\alpha$ and because $I$ is $\nu-invariant$:

\begin{align*}
    \{(X_e^\alpha) \in I \} & = \bigcap_{n \geq N} \{L_n^\alpha \in r_n(I)\} \\
    & \overset{\nu-a.s.}{=} \bigcap_{n \geq N} \; \bigcup_{\tau \in Sym([n])} \{L_n^\alpha \in \tau (r_n(I))\}.
\end{align*}

\noindent Lastly, because $\{M_n = [x]_\sim\} = \{L_n \in [x]_\sim\}$ by definition of our unlabeled/labeled chains:

\begin{align*}
    &    = \bigcap_{n \geq N} \{M_n^\alpha \in [r_n(I)]_\sim \} \\
    & \in \sigma(M_N^\alpha, M_{N+1}^\alpha, ...).
\end{align*}
% 
\noindent As $N \in \N$ was arbitrary, this implies $\{(X^\alpha_e) \in I\}$ lies in the tail $\sigma$-field of our bridge process $(M_n^\alpha)$, up to $\nu$-null sets. By assumption of bridge extremality, this tail $\sigma$-field is trivial and thus $\P((X^\alpha_e) \in I) \in \{0,1\}$ as desired. Hence the array $(X^\alpha_e)$ is ergodic.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:relating_different_notions_of_embedding} (Embedding and Morphism Densities)]
These proofs are simple extensions of the proofs in \cite{Meliot2024} of analogous results in the theory of graphons. Firstly (i) is immediate as any embedding is automatically a weak embedding. For (ii) note that if $\sigma \in Sym(\N)$ is such that $x \lesssim \sigma(y)$ then there is precisely one $x' \geq x$ so that $x' \leq y$, namely $x'$ is the array $\sigma(y)$ restricted to \red{the same size as $x$}. For (iii) if for each $e$ where $x_e = 0$ we let $A_e$ be the set of weak embeddings of $x' = x + \delta_e$ into $y$ then:
%
\[\{\text{embeddings x into $y$}\} = \{\text{weak embeddings $x$ into $y$}\} \setminus \bigcup_{e \;:\; x_e=0} A_e.\]
%
\noindent Then the formula follows from standard inclusion-exclusion. The proofs of (iv) and (v) can be found in the Appendix \ref{sec:proof_appendix}.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:equivalence_of_dm_and_density_convergence} (Doob-Martin Convergence and Embedding Densities)]
The equivalence of (ii) and (iii) follows immediately from Lemma \ref{lem:relating_different_notions_of_embedding}. The equivalence of (i) is almost an immediate result of exchangeability which implies that:

\[\P(L_n = x \; |\; M_n = [x]_\sim) = \frac{\P(L_n = x)}{\P(M_n = [x]_\sim) }=\frac{1}{| \; [x]_\sim \; |}.\]

\noindent As $L_n$ has deterministic backwards dynamics as we already noted above, this implies the unlabeled chain $M_n$ is uniformly backwards exchangeable where the labeling just associates $[x]_\sim$ to one element of the equivalence class. Hence, by \red{Theorem 3.1 of the other paper} we can see the appropriate notion of embedding density is:

\[\frac{| \; [y]_\sim \cap \pi_n^{-1}([x]_\sim) \; |}{|\; [y]_\sim \; |} \quad x \in E_n, y \in E_{n+m},\]

\noindent where $\pi_n^{-1}( [x]_\sim)$ is the collection of all arrays of the same size as $y$ that can be made by extending some array $\tx \in [x]_\sim$. To see that this equals \ref{def:embedding_density} note if $y =\sigma(y)$ for any $\sigma \in Sym([n+m])$ \red{then the numerator and denominator of the above equation both scale by same amount; in case this is true for no $\sigma$ we have immediate equality}.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:integral_representations_of_embedding_densities} (Embedding Densities of Representation Functions)]
First we prove the expectation formulas, starting with $emb^W$. Let $\phi : [k] \to [n]$ be an injective function. Then
    \begin{align*}
        \mathbb P(\text{$\phi$ is an array embedding}) &= \int_{[0,1]^{2^n}} \prod_{S \in \mathbf 1(F)} f(x_{(\phi(S))}) \, d\lambda_{2^n}
        \intertext{By the injectivity of $\phi$ we can apply a change of coordinates,}
        &= \int_{[0,1]^{2^n}} \prod_{S \in \mathbf 1(F)} f(x_{(S)}) \, d\lambda_{2^n} \\
        \intertext{After integrating out all the superfluous coordinates,}
        &= emb^W(F,f).
    \end{align*}
    So, writing $emb^W(F,X_n(f))$ in terms of the indicator functions $\mathbf 1_{\{\text{$\varphi$ is an array embedding}\}}$,
    \begin{equation*}
        \mathbb E[emb^W(F,X_n(f))] = emb^W(F,f).
    \end{equation*}

    \noindent The proof for $emb$ is identical. 
    
    For the concentration results, we follow exactly the proof of Theorem 2.5 in \cite{lovasz2004limits_of_graphs}. Namely, for $m \leq n$ let $X_m$ denote the (random) array induced by restricting $X_n(f)$ to subsets $e \subseteq [m]^{(d)}$. Since $\sigma(X_m) \subseteq \sigma(X_{m+1})$ since the former is a subarray of the latter, this can be used to define a Doob martingale:
    %
	    \[B_m := \E (emb^W(F, X_n(f)) \; | \; X_m).\]
    
    \noindent Bounding the increments using a bound on the number of injective functions $\phi:[k] \to [n]$ and applying Azuma's inequality as in \cite{lovasz2004limits_of_graphs} gives the desired inequalities. Then apply an identical procedure for $emb$. 

    The convergence then follows immediately from Borel-Cantelli. Lemma \ref{lem:relating_different_notions_of_embedding} (iv) and (v) allow the convergence to be extended to the morphism densities as well.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:metric_stronger_than_embeddings} (Cut Metric Controls Embedding Densities)]
The second inequality follows just by optimizing over the last coordinate, so we only need to prove the first inequality.
    \begin{align*}
        \text{LHS} &= \left | \int_{[0,1]^{2^k}} \prod_{S \in \mathbf 1(F)} f(x_{(S)}) - \prod_{S \in \mathbf 1(F)} f'(x_{(S)}) \, d\lambda_{2^k} \right |
        \intertext{Enumerate $\mathbf 1(F) = \{ S_1,S_2,\dots,S_\ell\}$. Using the triangle inequality, interpolating between the products by changing one factor at a time, we get}
        &\leq \sum_{m=1}^\ell \left| \int_{[0,1]^{2^k}} \left( \prod_{i=1}^{m-1} f(x_{(S_i)})  \right) (f(x_{(S_m)}) - f'(x_{(S_m)})) \left( \prod_{i=m+1}^{\ell} f'(x_{(S_i)})  \right) \, d\lambda_{2^k} \right |
        \intertext{To bound each of the $\ell$ terms, first integrate out all $2^k - 2^d$ variables not included in $x_{(S_m)}$. What remains after integrating the two product terms over those variables is a function $g_m$ of coordinates $\{x_A : A \subsetneq S_m\}$, so}
        &\leq |\mathbf 1(F)| \sup_{0 \leq g \leq 1} \left| \int_{[0,1]^{2^d-1}} \int_0^1 g(x_{-S}) (f(x_{-S},x_S) - f'(x_{-S},x_S)) \, dx_S \, d\lambda_{2^d-1}(x_{-S}) \right| \\
        &= |\mathbf 1(F)| \delta(f,f').\qedhere
    \end{align*}
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:automorphism_formula_for_sampling_probabilities} (Sampling Probability Formula)]
From Lemma~\ref{lem:relating_different_notions_of_embedding}, we have that
    \begin{align*}
        emb^W(x,f) &= \mathbb E[emb^W(x,A_n)] \\
        &= \sum_{|y| = n} \mathbb P(A_n = y) emb^W(x,y) \\
        &= \sum_{|y| = n} \sum_{x' \geq x} \mathbb P(A_n = y) emb(x',y) \\
        \intertext{Since $x$ and $y$ are both of size $n$, an embedding is the same as an automorphism of the finite array. In particular, $emb(x',y) = 0$ if $x',y$ are not isomorphic, and $emb(x',y) = |\text{Aut}(x')|/n!$ if they are isomorphic.}
        &= \sum_{x' \geq x} \sum_{y \text{ iso.\ to } x'} \mathbb P(A_n = x') \frac{|\text{Aut}(x')|}{n!} \\
        &= \sum_{x' \geq x} \mathbb P(A_n = x') \frac{|\text{Aut}(x')|^2}{n!}.
    \end{align*}
    By M\"obius inversion, we get
    $$\frac{|\text{Aut}(x)|^2}{n!} \mathbb P(A_n = x) = \sum_{x' \geq x} (-1)^{|1_x| - |1_{x'}|} emb^W(x',f),$$
    and the result follows.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:tv_bound_between_samples} (Total Variation Bound for Samples)]
By Lemma~\ref{lem:automorphism_formula_for_sampling_probabilities}, for any fixed array $x$ of size $n$,
    \begin{align*}
        |\mathbb P(A_n = x) - \mathbb P(A_n' = x)| &\leq \frac{n!}{|\operatorname{Aut}(x)|^2} \sum_{x' \geq x} |emb^W(x\red{'},f) - emb^W(x',f')| \\
        &\leq n! \sum_{x' \geq x} |emb^W(x\red{'},f) - emb^W(x',f')|.
    \end{align*}
    Summing over all $x$, we get
    $$\sum_{|x| = n} |\mathbb P(A_n = x) - \mathbb P(A_n' = x)| \leq \red{n!}\sum_{|x| = n} \sum_{x' \geq x} |emb^W(x,f) - emb^W(x',f')|.$$
    The left hand size is twice the total variation distance between the laws of $A_n,A_n'$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:coupling_of_samples} (Coupling of Samples)]
If we can show that
    \[
    \| \text{Law}(A_n) - \text{Law}(A_n') \|_{\text{TV}}
    \leq \frac{n!}{2} 3^{\binom{n}{d}}
    \max_{|x| = n} |emb^W(x,f) - emb^W(x,f')|,
    \]
    then there exists a coupling of $A_n,A_n'$ such that $\mathbb P(A_n \neq A_n') \leq []$. Since $\delta_s(f_n, f_n') = 0$ when $A_n = A_n'$ and $\delta_s(f_n, f_n') \leq 1$, this will give
    \begin{align*}
        \mathbb E[\delta_s(f_n,f_n')] &= \mathbb E[\delta_s(f_n,f_n') \mathbf 1_{\{A_n \neq A_n'\}}] \\
        &\leq \mathbb E[ \mathbf 1_{\{A_n \neq A_n'\}}] \\
        &\leq \frac{n!}{2} 3^{\binom{n}{d}} \max_{|x| = n} |emb^W(x,f) - emb^W(x,f')|.
    \end{align*}
    
    To prove the total variation bound, we use Corollary~\ref{cor:tv_bound_between_samples}:%However, we will apply the bound to $g_n := \frac{1}{2}(1 + f)$ and $g_n' := \frac{1}{2}(1 + f')$, noting that $\delta_s(g_n,g_n') = \frac{1}{2} \delta_s(f_n,f_n')$.
    \begin{align*}
        \| \operatorname{Law}(A_n) - \operatorname{Law}(A_n') \|_{\operatorname{TV}} &\leq \frac{n!}{2} \sum_{|x| = n} \sum_{x' \geq x} |emb^W(x',f) - emb^W(x',f')| \\
        &\leq \frac{n!}{2} 3^{\binom{n}{d}} \max_{|x| = n} |emb^W(x,f) - emb^W(x,f')|.
    \end{align*}
     because for each of the ${n \choose d}$ entries of our arrays $x', x$ we have three choices that ensure $x' \geq x$: $(x_e, x_e') \in \{(0,0), (0,1), (1,1)\}$. \qedhere
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:conv_embed_implies_conv_cut_metric} (Convergence via the Cut Metric)]
(iii) $\implies$ (ii): This follows by definition, as $\delta(g_n,g) \leq \delta_s(g_n,g)$.

    (ii) $\implies$ (i): This follows from Lemma~\ref{lem:metric_stronger_than_embeddings}.

    (i) $\implies$ (iii): For any given $N$, let $A_N^n,A_N'$ be random arrays of size \red{$N$}, sampled from $g_n,g$, respectively, and let $f_N^n,f_N'$ be the associated sampling functions. The triangle inequality gives
    \begin{align*}
        \delta_s(g_n,g) &\leq \E[\delta_s(g_n,f_N^n)] + \E[\delta_s(f_N^n,f_N')] + \E[\delta_s(f_N',g)],
        \intertext{where we have coupled $f_N^n,f_N'$ as in Lemma~\ref{lem:coupling_of_samples}. Using Lemmas~\ref{lem:finite_functions_converge_to_overall_sampling_functions} and \ref{lem:coupling_of_samples}, we get}
        &\leq \frac{10}{\sqrt{\log_2 N}} + \frac{N!}{2} 3^{\binom{N}{d}} \max_{|x| = N} |emb^W(x,g_n) - emb^W(x,g)|.
    \end{align*}
    Letting $n \to \infty$, since there are only finitely many possible arrays of size $N$, we get
    $$\limsup_{n \to \infty} \delta_s(g_n,g) \leq \frac{10}{\sqrt{\log_2 N}}.$$
    Letting $N \to \infty$ completes the proof.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:approximation_by_step_fn} (Step-Function Approximation)]
Let $f_\cP := \E [f \; | \cG_\cP]$
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:finite_functions_converge_to_overall_sampling_functions} (Approximation by Sampled Arrays)]
Let $\varepsilon > 0$, the value of which will be determined later. First, we will approximate $f$ by $g := \E[f \mid \mathcal G_{\mathcal P}]$, where $\mathcal P$ is as given in Lemma~\ref{lem:approximation_by_step_fn}. We also approximate $f_n$ by $g_n$, where $g_n$ is the (random) sampling function corresponding to a random array of size $n$, sampled from $g$.Then
    \begin{align*}
        \mathbb E[\delta_s(f_n,f)] &\leq \mathbb E[\delta_s(f_n,g_n)] + \E[\delta_s(g_n,g)] + \delta_s(g,f) \\
        &\leq \mathbb E[\delta_s(f_n,g_n)] + \E[\delta_s(g_n,g)] + \varepsilon.
    \end{align*}

    To control $\delta_s(g_n,g)$, we will find a version of $g_n$ for which the partition lines up well with $\mathcal G_{\mathcal P}$. Writing $\mathbb P = \{I_1,\dots, I_q \}$, let $U_1,\dots,U_n \overset{\operatorname{iid}}{\sim} \text{Unif}[0,1]$, and let $N_i = | \{ j : U_j \in I_i\}|$, so that $N_i \sim \text{Binomial}(n,\lambda(I_i))$. Then define a version of $g_n$ as follows: Partition $[0,1]$ into $J_1,\dots,J_q$ in such a way that $\lambda(J_i) = N_i/n$ (note that $N_1 + \cdots + N_q = n$) and $J_i \cap I_i = \min(\lambda(I_i),\lambda(J_i))$; that is, we line up the empirical partition as well as we can with $\mathcal P$. Then, letting $g(I_{i_1} \times \cdots \times I_{i_d})$ denote the value that $g$ takes on $I_{i_1} \times \cdots \times I_{i_d}$, we set $\widetilde g_n(x_1,\dots,x_d)$ to equal the value $g$ takes on $I_{i_1} \times \cdots \times I_{i_d}$ when $x_1 \in J_{i_1},\dots,x_d \in J_{i_d}$. 
    Using this version of $g_n$,
    \begin{align*}
        \delta_s(g_n,g) &\overset{d}{=} \delta_s(\widetilde g_n,g) \\
        &\leq \int_{[0,1]^d} |\widetilde g_n-g| \, d\lambda_d \\
        \intertext{These functions both take values in $[0,1]$, so we can upper bound the integral by the measure of the set where the functions disagree.}
        &\leq \lambda(\{ x : \widetilde g_n(x) \neq g(x) \}) \\
        &\leq 1 - \left( \sum_{i=1}^q \min( \lambda(I_i),\lambda(J_i)) \right)^d \\
        \intertext{Using $1-x^d = (1-x)(1 + x + \cdots +x^{d-1}) \leq d(1-x)$ for $0 \leq x \leq 1$,}
        &\leq d \left( 1 - \sum_{i=1}^q \min( \lambda(I_i),\lambda(J_i)) \right) \\
        \intertext{Using the standard total variation identity for finite partitions,}
        &= \frac{d}{2} \sum_{i=1}^q \left| \lambda(I_i) - \lambda(J_i) \right| \\
        &\leq \frac{d\sqrt q}{2} \sqrt{\sum_{i=1}^q (\lambda(I_i) - \lambda(J_i))^2}.
    \end{align*}
    Taking the expectation of both sides of the inequality, we get
    \begin{align*}
        \E[\delta_s(g_n,g)] &\leq \frac{d \sqrt q}{2} \E \left[ \sqrt{\sum_{i=1}^q \left(\lambda(I_i) - \frac{N_i}{n}\right)^2} \right] \\
        &\leq \frac{d \sqrt q}{2} \sqrt{\E \left[ \sum_{i=1}^q \left(\lambda(I_i) - \frac{N_i}{n}\right)^2 \right]} \\
        &= \frac{d \sqrt q}{2} \sqrt{\sum_{i=1}^q \text{Var}(N_i/n)} \\
        &\leq \frac{d \sqrt q}{2 n} \sqrt{\sum_{i=1}^q n \lambda(I_i)} \\
        &= \frac{d}{2} \sqrt{\frac{q}{n}}.
    \end{align*}

    To bound $\delta_s(f_n,g_n)$, we combine Lemma~\ref{lem:coupling_of_samples} with Lemma~\ref{lem:metric_stronger_than_embeddings}:
    \begin{align*}
        \E[\delta_s(f_n,g_n)] &\leq \frac{n!}{2} 3^{\binom{n}{d}} \max_{|x|=n}  |emb^W(x,f) - emb^W(x,g)| \\
        &\leq \frac{n!}{2} 3^{\binom{n}{d}} \binom{n}{d} \delta_s(f,g) \\
        &\leq \frac{n!}{2} 3^{\binom{n}{d}}  \binom{n}{d} \varepsilon.
    \end{align*}
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:weak_embeddings_for_nonbinary_arrays} (Weak Embeddings for Nonbinary Arrays)]
The proof is analogous to that of Lemmas \ref{lem:relating_different_notions_of_embedding} and \ref{lem:equivalence_of_dm_and_density_convergence}. Namely, for (i) note that if $\phi$ is a C-weak embedding there is exactly one choice of $x' \geq_C x$ so that $\phi$ is an embedding of $x'$ into $y$. Namely, $x'_e := y_{\phi(e)}$ whenever $x_e \notin C$. 

    For (ii), note any C-weak embedding that is not an embedding must have $y_{\phi(e)} \neq x_e$ for some $x_e \notin C$. Thus:
    %
\[emb(x,y) = emb^C(x,y) \setminus \bigcup_{e : \; x_e \notin C} \; \bigcup_{t \neq x_e} \{ \phi \in emb^C(x,y) \; : \; y_{\phi(e)} = t \}.\]
    %
    \noindent Let $A_{e,t}$ denote the set inside the union. Note
        %
        \[emb^{C \cup \{t\}}(x',y) \subseteq A_{e,t} \subseteq  emb^C(x',y) \subseteq emb^C(x,y),\]

        \noindent where $x'$ is obtained from $x$ by replacing $x_e$ by $t$. This is because the first term requires $y_{\phi(e)} = t$ for all $x'_e = t$, the second requires it for only our special index $e$, and the last term does not require it at all. If $t \in C$ the first three terms are equal but the last two are not, and if $t \notin C$ then the last two terms are equal but the first three are not. More generally for distinct $e_i$: 
        %
        \begin{equation}
                emb^{C \cup \{t_1, ..., t_m\}}(x',y) \subseteq \bigcap_{i=1}^m A_{e_i, t_i} \subseteq emb^C(x',y) \subseteq emb^C(x,y)
                \label{eqn:A_ei,ti_bounds}
        \end{equation}
 
        %
        \noindent where $x'$ is obtained from $x$ by replacing $x_{e_i}$ by $t_i$.  By inclusion-exclusion:
    %
    \begin{align*}
        |\induced(x,y)| & = \sum_{m \geq 0}\sum_{\substack{e_1,..., e_m \\ distinct \\ t_1,..., t_m}} (-1)^m\bigg| \bigcap_{i=1}^m A_{e_i, t_i} \bigg| \\
        %
        & = \red{\sum_{x' \geq_C x} (-1)^{d_{ham}(x,x')} \; |A_{x,x'}|}, 
    \end{align*}
%
    \noindent where $d_{ham}(x,x')$ is the hamming distance, the number of entries $x,x'$ differ, and:
    %
    \[A_{x,x'} = \bigcap_{e \; : \; x_e \neq x'_e} \; A_{e, x'_e},\]
    %
    The desired result then follows from applying inequality \eqref{eqn:A_ei,ti_bounds}.
    
    For (iii), note that $t \neq x_e = s$ by assumption so we always fall into the case where $t \in C$ and thus $A_{e,t} = emb^C(x',y )$. Now we can just apply inclusion-exclusion as in the previous part.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:reducing_nonbinary_conv_to_binary} (Binary Projections of Nonbinary Arrays)]
% proof formula
    Suppose $\phi:|x| \to |y|$ is injective. Then by construction of our one-hot encoding $x_e = y_{\phi(e)}$ if and only if $g_s(x_e) = g_s(y_{\phi(e)})$ for all $s \in S$, and hence the first statement follows. 
    
    % proof conv implies convergence of projections
     Now suppose that $emb(\cdot, y_N)$ converges and that $S = \{0,...., n\}$. If $A = S$ then $g_A(y_N)$ is the array of all ones for every $N$, so the result is trivial. Thus without loss of generality assume $0 \notin A$. By Lemma \ref{lem:equivalence_of_dm_and_density_convergence}, it suffices to show $emb^W(b, g_A(y_N))$ converges for an arbitrary binary array $b$. Define a set of finite arrays of the same size of $b$ by:
    %
    \[\cX^b_A := \{x \in S^{[|b|]^{(d)}} \; : \; x_e \in A \text{ if } b_e = 1 \text{ and } x_e = 0 \text{ otherwise}\} \subset g_A^{-1}(b).\]
    %
    Namely, $\cX^b_A$ is $g_A^{-1}(b)$ restricted to arrays with entries only in $A \cup \{0\}$. We aim to show:
    %
    \begin{equation}
        emb^W(b, g_A(y_N)) = \bigsqcup_{x \in \cX^b_A} emb^{S \setminus \{0\}}(x, y_N)
        \label{eqn:relating_embeddings_into_yn_to_those_into_gA(yN)}
    \end{equation}
    
    %
    To see why the latter sets are disjoint, suppose that $\phi \in emb^{S \setminus \{0\}}(x, y_N) \cap emb^{S \setminus \{0\}}(x', y_N)$. Note by the definition of $\cX^b_A$ that $x_e = 0$ iff $b_e=0$ iff $x_e' = 0$. Furthermore, for each $e$ where $x_e,x_e' \neq 0$ we have $x_e = (y_N)_{\phi(e)} = x'_e$. Thus $x = x'$. 
    
    For the forward inclusion, suppose $\phi \in emb(b, g_A(y_N))$. Define a finite array $x$ by:

	    \[x_e := \begin{cases}
	        (y_N)_{\phi(e)} & b_e = 1 \\
	        0 & b_e = 0 \\
	    \end{cases}.\]
    %
    \noindent If $b_e = 1$ then $g_A(y_N)_{\phi(e)} = 1$ and thus $(y_N)_{\phi(e)} \in A$. Thus $x \in \cX^b_A$, and by construction $\phi \in emb^{S \setminus \{0\}}(x,y_N)$. 

    For the reverse inclusion, suppose $\phi \in emb^{S \setminus \{0\}}(x,y_N)$ for some $x \in \cX^b_A$. Then for each $e$ with $x_e \neq 0$ we have $(y_N)_{\phi(e)} = x_e \in A$.  Since $x_e \neq 0$ if and only if $b_e = 1$ this implies $b_e = g_A(y_N)_{\phi(e)}$ as desired. Thus $\phi \in emb^W(b, g_A(y_N))$. This completes the proof of equation \eqref{eqn:relating_embeddings_into_yn_to_those_into_gA(yN)}. 
    
    Dividing both sides of \eqref{eqn:relating_embeddings_into_yn_to_those_into_gA(yN)} by the number of injective functions $\phi: |x| \to |y_N|$ yields:
    %
\begin{equation}
      emb^W(b, g_A(y_N)) = \sum_{x \in \cX_A^b} emb^{S \setminus \{0\}}(x, y_N).
\end{equation}
    %
     The result then follows by applying Lemmas \ref{lem:weak_embeddings_for_nonbinary_arrays} and \ref{lem:equivalence_of_dm_and_density_convergence}.\\


    % proof conv of projections implies convergence.
    \noindent \red{TODO} Conversely, suppose $emb(\cdot, \; g_A(y_N))$ converges for all $A \subseteq S$ and let $x$ be a $S$-valued array. 

    \zack{Choosing $A = \{s\}$}
    
    \zack{Hopefully we can just reverse equation \eqref{eqn:relating_embeddings_into_yn_to_those_into_gA(yN)} to show that $emb^{S \setminus \{0\}}$ converges and then apply lemma \ref{lem:weak_embeddings_for_nonbinary_arrays}}
\end{proof}

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 
\section{TODO}
\label{sec:exchangeable_array_todo}


\begin{enumerate}
    \item \red{DONE!} Understand what infinite bridges of exchangeable array processes look like. I suspect they themselves can be framed as some kind of exchangeable array process.

    \item \red{DONE!} Understand what the extremal infinite bridges of an exchangeable array process looks like, and hence the Doob Martin boundary of such chains. Following ideas in \cite{DM_Remy} and the ballot chain paper, I imagine these are just the infinite bridges that are ergodic in the sense of exchangeable arrays. In terms of their representation functions $f$ I feel these ought to be the ones where $f$ depends trivially on the first coordinate. For example, with our graph sequences these are the ones corresponding to graphons $f(u,x,y,z) := \ind \{z \leq w(x,y)\}$. 


    \item \red{DONE!} Determine how the Doob-Martin topology relates to the embedding density, which is a natural notion in both exchangeable array processes and our backwards exchangeable chains. Namely, we would like to say you converge in the Doob-Martin topology iff the embedding density converges for every fixed input (e.g. for graphons this is the morphism density).

    \begin{itemize}
        \item \red{DONE}: Develop a notion of embedding density for our candidate limits: the representation functions of our exchangeable arrays. Analogous to graphons, $t(F,f)$ ought to have the property that it equals the average embedding density of $F$ into a finite array sampled from $f$.
    \end{itemize}
    
    
    \item Can we metrize this topology like graphons manage to do using the cut metric? Does the cut metric make sense for \textit{any} array (beyond just the adjacency matrices of our graph chains; possibly we have to restrict to binary arrays?)

    \begin{itemize}
        \item Namely, it seems like we want to create some sort of metric on the sampling representations $f$ of the exchangeable array. Since the $f$ from the sampling representation is not unique per say we will have to define some sort of equivalence classes, analogous to how graphons are really only defined up to Lebesgue isomorphisms. Steve says a similar ``unique up to isomorphism'' statement is true for exchangeable arrays representations $f$ and that the details are in Kallenberg (Theorem 7.28 in his book on Symmetry). 

        \item \textbf{GOAL:} relate the cut metric to the sampling representation for graphons we deduced above. Maybe this can give us an idea of how to define the metric in a more general setting. It may depend on what the state space $E$ is of the array of course, so maybe we assume this is some kind of metric space in its own right.

        \item \textbf{GOAL:} Show that convergence in this extended cut metric is equivalent to convergence of the embedding densities (and hence also in the Doob-Martin topology). 

        \begin{itemize}
            \item Find a suitable notion of embedding a finite array $x$ into our space of representation functions, analogous to how finite graphs are embedded in the space of graphons just by discretizing the adjacency matrix. This should be done in such a way so that the embedding density $t(F, x)$ is the same whether we view $x$ as a finite array or as a representation function. 
        \end{itemize}

        \item This would be really cool if we could do this, as this could feasibly give us a way of comparing the exchangeable arrays corresponding to completely different combinatorial processes.

        \daniel{This is an interesting idea. Maybe this would depend on the realization as an exchangeable array.}
    \end{itemize}


    \item Once we generalize the cut metric, can we extend any pre-existing theory of graphons to our new exchangeable array processes (and hence other combinatorial stochastic processes?). e.g. Can we extend theory of large deviations of random graphs to large deviations of other combinatorial structures? Can we interpret certain graph properties like the existence of independent sets, cliques, etc in these new settings? This aim may seem a bit vague, but we have a lot of pre-existing theory of graphons to work with. Review how the cut metric and graphon convergence has been used to prove results in extremal graph theory or spectral graph theory, and try to extend some of these results to other combinatorial stochastic processes.  

    \item Try to relate the sampling representation of exchangeable arrays to some of the sampling representations we have so far in Doob-martin theory (e.g. Remy's chain, ballot chain, Polya's urn, etc.). Can we say something more in general on what kind of measures are induced?

    \item It is easy to see every exchangeable array process is backwards exchangeable. Try to determine the proper constraints for the converse to hold: which backwards exchangeable chains can be encoded as an exchangeable array process, and try to find a general scheme for doing so. So far all the examples we have seen above fall into both frameworks, so the class of such chains must be fairly large.

    \item Not every combinatorial stochastic process can be encoded by the ``k-wise'' relationships between its parts. Daniel gave the example of the hypergraph chain. This seems to require a theory of ``infinite dimensional'' exchangeable arrays. One avenue to explore would be how to make this rigorous, and if there is a similar theory of representation functions for such arrays.

    \begin{itemize}
        \item \zack{As far as I can tell, Kallenberg's theory is not restricted to arrays of a fixed dimension $d$, unlike what is stated in Tim Allen's exchangeable arrays notes. Theorem 7.22 In \cite{KallenbergProbabilisticSymmetries} is stated for any random array indexed by ALL subsets of $\N$. Tim Austin has some notes in a paper called \href{https://arxiv.org/abs/1302.2116}{Exchangeable Random Measures} which makes it a bit more clear what the representation function looks like (Theorem 1.2). It just seems like you require one middle-symmetric function for each dimension, and you reuse the randomness from the uniform random variables of the lower dimensions.}
    \end{itemize}

    \item What we have done above extends the theory of graphons of dense graphs to some notion of ``dense'' combinatorial stochastic processes. The \textit{graphex model} and the notion of \textit{sampling convergence} discussed in \cite{Borgs_2019_Sampling_convergence_sparse_graphs} extends these notions to sparse graphs. We could try to do some of the same if we want.

    \item One sort of vague idea I had was it seems is that we often design these Markov chains in order to generate the combinatorial objects uniformly. The relation to exchangeable arrays comes from the fact that the dynamics of our chain has some sort of invariant that is perserved as the chain progresses (e.g. ancestry in Remy chain), use that to encode my process as an exchangeable array, and develop my limit theory from there. Could we in some sense work backwards? Like if I had some sort of invariant, could I define an array using it. Such arrays already have a natural Markov chain associated to it which we talk about. Could we say this generates the combinatorial objects uniformly, or maybe at least that some bridge of this chain can do this? Under what conditions does such a bridge exist?

    \item The paper \cite{forman2017representationexchangeablehierarchiessampling} gives a representation of exchangeable hierarchies in terms of exchangeable arrays. It would be interesting to explore what the associated markov chain is, and what its notion of embedding would be. They also in a sense characterize the extreme points of the limiting object as those exchangebale hierarchies on $\N$ which are independently generated. Can we deduce this immediately from our characterization of the extremal exchangeable arrays (the ergodic ones; specifically the equivalent independence condition stated in theorem \ref{thm:boundary_is_ergodic_arrays}). 

    \item In the papers \cite{forman2017representationexchangeablehierarchiessampling} and \cite{DM_Remy}, they encode their exchangeable processes as exchangeable arrays, and use these to build real trees from which the processes are encoded as sampling from these real trees. The former paper relies on a simple stick-breaking construction, and I wonder if this would be possible to do in general. This would allow us to view all exchangeable array processses as some sort of sampling process from a real tree. The array they use is:

	    \[X_{ij} = 1 - \lim_{n \to \infty} \frac{\# \text{leaves below mrca of i,j}}{n}.\]
    
    \red{similar to the distance function steve uses in his Remy chain paper. If we define a 3-dimensional binary array $X_{ijk} = \ind\{\text{k is below mrca(i,j)}\}$ then this basically corresponds to taking an average of the last coordinate. Maybe we could do the same in general for a 3D binary array, or maybe any binary array?}

    \item Is there a way of encoding the randomness in exchangeable array processes to ensure the underlying combinatorial objects
          are generated uniformly? For example, in the Remy chain, \cite{DM_Remy} the underlying algorithm generates trees uniformly 
          in an iterative fashion, and \cite{Marckert_generating_d-ary} extends this to d-ary trees.
\end{enumerate}
