%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass{ucbthesis}
\usepackage{rotating} % provides sidewaystable and sidewaysfigure

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\addtolength{\abovecaptionskip}{\baselineskip}

\addbibresource{references.bib}

\hyphenation{mar-gin-al-ia}
\hyphenation{bra-va-do}

\begin{document}

% Declarations for Front Matter

\title{Thesis Title}
\author{Zachary McNulty}
\degreesemester{Summer}
\degreeyear{2026}
\degree{Doctor of Philosophy}
\chair{Professor Steven N. Evans}
\othermembers{Professor Shirshendu Ganguly \\
   Professor Fraydoun Rezakhanlou}
% For a co-chair who is subordinate to the \chair listed above
% \cochair{Professor Benedict Francis Pope}
% For two co-chairs of equal standing (do not use \chair with this one)
% \cochairs{Professor Richard Francis Sony}{Professor Benedict Francis Pope}
\numberofmembers{3}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Applied Mathematics}
% Designated Emphasis -- this is optional, and rare
% \emphasis{Colloidal Telemetry}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional (default is Berkeley)
% \campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

% \begin{dedication}
% \null\vfil
% \begin{center}
% Dedication.
% \end{center}
% \vfil\null
% \end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\setpnumwidth{2em}
{\sloppy\tableofcontents}
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}

   I would like to express my gratitude to the many individuals who have supported
me throughout my Ph.D. journey.

First and foremost, I would like to thank my advisor, Professor Steven Evans, for his patience, guidance, and support throughout 
these many challenging years. His kind words, gentle encouragement, and insightful feedback
have been invaluable in shaping my research and guiding me through this tumultous process. 
I am grateful for the opportunity to work with him and for his mentorship. Thanks to my committee members, 
Professors Shirshendu Ganguly and Fraydoun Rezakhanlou, for their insightful feedback and encouragement,
for starting me off at the beginning of my Ph.D. journey, and now for seeing it to a close. Thanks to
my colloborator, Daniel Raban, for his help on part of this thesis, and for his friendship along the way.

I would like to thank my partner Julia Kim for her unwavering love and encouragement. She has been a constant source of strength, believing in me
on the many days I did not believe in myself, and for that I am eternally grateful. I am not sure
how I would have completed this journey without her by my side. To my cat, Zack Junior, for being cute, and spending many
afternoons in his cozy cat bed by my side.

Last, but not least, I would like to thank my family and friends for their unwavering support and encouragement. 
To the many members of the math department soccer (go Possums!) and frisbee team, and our regular group 
of spire slayers and bird collectors, for keeping me sane throughout this process, and for helping me find 
a sense of belonging in the department. To Eric Jankowski, Andrew Scharf, Carlos Esparza, Connor Halleck-Dube,  
Tom Schang, Lucy Horowitz, Spencer Taylor, Ely Sandine, and the other members of my cohort 
Rose Lopez, John Nolan. To Zack Stier and our many afternoons of basketball. I wish you all the best of luck on the 
next step of your journey.

\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\part{Completed Projects}
\graphicspath{{chapters/bipartition_covers/figures/}}
\input{chapters/bipartition_covers/chapter}

\graphicspath{{chapters/self_balancing_sampling/}}
\input{chapters/self_balancing_sampling/chapter}

\part{Developing Projects}

\graphicspath{{chapters/sym_max_fair_values/}}
\input{chapters/sym_max_fair_values/chapter}

\graphicspath{{chapters/canby/}}
\input{chapters/canby/chapter}

\graphicspath{{chapters/entropy_regularized_opt_control/}}
\input{chapters/entropy_regularized_opt_control/chapter}

\graphicspath{{chapters/exch_array_processes/}}
\input{chapters/exch_array_processes/chapter}



\printbibliography

% \appendix
% \chapter{More Monticello Candidates}

\end{document}
