\chapter{Inference under the Canby Model}

\begin{center}
    \textbf{Abstract}
\end{center}

\noindent This chapter studies phylogenetic inference under the Canby model, a stochastic model for
polymorphic linguistic characters. We focus on the no-borrowing setting and develop basic probabilistic
properties of the model, with particular attention to how non-homoplastic states generate quartet support
and candidate bipartitions. These results motivate ASTRAL-based inference procedures and provide preliminary
sample-complexity bounds for recovering the underlying language phylogeny from independently evolving
characters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

In phylogenetics, we have a collection of objects whose ancestral history we hope to infer
using only modern-day information. This shared history is summarized by a \textit{phylogenetic tree},
a graphical tree whose leaves are labeled by these objects. Classically, these objects represented different biological species,
the relevant data was their DNA, and the phylogenetic tree encoded the sequence of
speciation events that occurred over time to produce the species we see today. Nonetheless, the tools of phylogenetics are far 
more broadly applicable: they have been used in medicine to better understand the 
progression of cancer~\cite{somarelli2017phylooncology}, in astrophysics as a hierarchical clustering procedure for astral
 bodies~\cite{fraixburnet2015multivariate}, and in terrane analysis to study the geological history of specific 
 regions~\cite{michaux2018cladistic}. 

In this chapter, we study a specific application to linguistics, where the objects of interest are different 
languages and the underlying data are \textit{linguistic characters}. These characters encode features 
that vary across languages and are commonly divided into three broad categories: \textit{phonological},
 \textit{morphological}, and \textit{lexical}. 
 
 Phonological characters describe sound patterns, 
 such as whether a language distinguishes tones or permits consonant clusters. Morphological characters 
 describe grammatical structure, such as how past tense or plurality is expressed. Lexical characters describe the word 
 forms used to express specific ideas, often by grouping languages according to whether their words for 
  common concepts such as ``fire,'' or ``mother'' descend from a common historical source. Lexical characters are 
  particularly interesting as they can be multi-state---several different words can simultaneously share 
  the same meaning---a property known as \textit{polymorphism}.  We are specifically interested in a model known as the \textit{Canby et al. model}, introduced in 
\cite{polymorphism_linguistic_phylogenetics}, which explicitly models this polymorphism.

%%%%%%%%%%%%%%%%%%%%
% Canby Model
%%%%%%%%%%%%%%%%%%%%
\subsection{The Canby et al. Model}

The key feature of the Canby model is that it allows for polymorphism: a given character $c$ may take on multiple different
states simultaenously. To differentiate between the set-valued collection of states a character currently takes on and the individual
states themselves, we call the former the \textit{macro-state} of the character and the latter its individual \textit{micro-states}. For
example, if the macro-state of a character $c$ at a given point of time is $\{\alpha, \beta\}$ then it currently exhibits two micro-states: $\alpha$ and  $\beta$. 
Under the Canby model, it is assumed that micro-states may be born or die as the character evolves along the tree, so that the current
\textit{number} of micro-states present in the current macro-state evolves according to a standard birth-death process, which we refer to as the \textit{cardinality process}.

More formally, suppose that $\cT^*$ is the underlying rooted phylogenetic tree of languages. Given some collection of characters $\cC$, 
associated to each character $c$ is a set $\Sigma^{(c)}$
of possible states, birth/death rates $\{\lambda_i^{(c)}\}, \{\mu_i^{(c)}\}$ controlling the cardinality process, a probability measure $\nu^{(c)}$ 
on $\Sigma^{(c)}$ for selecting new states, and a probability measure $\rho^{(c)}$ on the set ${\Sigma^{(c)} \choose *}$ of nonempty finite subsets of $\Sigma^{(c)}$ for 
determining the macro-state of the root. Let $[0, \ell]$ denote the unique path from the root of $\cT^*$ to some leaf $\ell \in \cT$. For $x \in [0, \ell]$ we assume the
macro-state $c(x)$ for the given character evolves down this path as follows
%
\begin{enumerate}[label=(\roman*)]
    \item The macro-state $c(0)$ of the root is sampled from $\rho^{(c)}$. 
    \item If $|c(x)| = i$, at rate $\lambda_i^{(c)}$ a new character is born, and at rate $\mu_i^{(c)}$ a character will die.
    \item If a death occurs, a state $\alpha \in c(x)$ is chosen uniformly at random to delete. If a birth occurs, a new state $\alpha \notin c(x)$ is chosen proportional to $\nu^{(c)}$, namely via
                \[\mu(A) := \frac{\nu^{(c)}(A \setminus c(x))}{\nu^{(c)}(\Sigma^{(c)} \setminus c(x))}\]
\end{enumerate}
%
When a branch-point $v$ of $\cT^*$ is reached, conditional on $c(v)$ the macro-state evolves independently down each child branch starting from
state $c(v)$. 

A state $\alpha \in \Sigma^{(c)}$ with $\nu^{(c)}(\{\alpha\}) > 0$ is called \textit{homoplastic}. These states
can evolve independently in different lineages, rather than from shared ancestry. We denote the set of homoplastic states
and non-homoplastic states by $\Sigma_H^{(c)}, \Sigma_N^{(c)}$ respectively. Homoplastic states in a sense complicate the 
inference process, as two languages can share the homoplastic trait without necessarily having any recent shared ancestry.
 
Throughout this paper, we will assume characters follow the \textit{immigration-death rates}
%
\begin{equation}
    \lambda_i^{(c)} = \lambda^{(c)}, \quad \mu_i^{(c)} = i \mu^{(c)} \cdot 1\{i > 1\}.
\end{equation}
%
\noindent for $i \geq 1$. Namely, the birth-rate is constant, and the death rate is proportional to the number of states,
except when only one state is present. These assumptions ensure that each character
has at least one state associated to it at any point in time and that the number of states cannot blow up in finite time.  

Since the upcoming inference procedure is based solely on non-homoplastic states, we also make the
following assumptions about $\nu^{(c)}$ and $\rho^{(c)}$.

\begin{assumption}[1]
    Non-homoplastic states occur sufficiently often. Namely
    %
    \begin{enumerate}[label=(\roman*)]
        \item There is positive probability that the macro-state at the root contains at
              least one micro-state that is not homoplastic
              %
                \[\rho^{(c)} \left(\left\{A \in {\Sigma^{(c)} \choose *} \; : \; A \cap \Sigma^{(c)}_N \neq \emptyset \right\}\right) > 0.\]
              %
        \item There is positive probability of adding a micro-state that is not homo-
                plastic to the current macro-state when there is a birth
                %
                \[\nu^{(c)}(\Sigma_N^{(c)}) > 0.\]
                %
    \end{enumerate}
\end{assumption}

The full Canby model additionally incorporates another mechanism of character transmission known
as \textit{borrowing} which allows characters to transfer laterally throughout the tree, analogous to
horizontal gene transfer in biological contexts. In this paper, we only consider the case where no borrowing
occurs, but interested readers can find further details in \cite{polymorphism_linguistic_phylogenetics}.


%%%%%%%%%%%%%%%%%%%%%%%%%
% ASTRAL
%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{ASTRAL}

To infer the underlying tree from many independently evolving characters, we adapt ideas from
quartet-based species-tree estimation. In phylogenetic inference, one often wants to combine information from many sources, such as genetic information across
a collection of genes or, in our case, linguistic characters. In \textit{consensus methods}, each source is
used to generate an estimate $g_i$ of the true phylogenetic tree $\cT^*$, and then these estimates $\{g_i\}_{i=1}^n$
are aggregated. In the phylogenetics literature, the trees $g_i$ are called the \textit{gene trees}, while $\cT^*$ is
referred to as the \textit{species tree}. We will occassionally utilize this vocabulary.

ASTRAL~\cite{mirarab2014astral} is one commonly used aggregation procedure. It scores candidate 
phylogenetic trees $\cT$ based on the number of \textit{quartets}, four-leaf induced subtrees, they share with the estimated trees $\{g_i\}$. Specifically,
it solves the optimization problem
%
\begin{equation}
    \hat{T} = \argmax_\cT  \sum_{q \in \cQ(\cT)} W(q).
    \label{eqn:astral_quartet_support_optimization_problem}
\end{equation}
%
\noindent where $W(q)$ is some meaure of the support for the quartet $q$. Commonly, it is just
$W(q) = \sum_{i=1}^n 1 \{q \in g_i\}$, the number of times the quartet appears in the gene trees. Later on, we will introduce
other measures of the quartet support specific to the Canby model.  
    
Since the set of phylogenetic trees is super-exponentially large, it is intractable to solve this 
optimization problem exactly for trees with more than 20 or so leaves. For larger trees, a
restricted search space is required. One possible strategy is to limit the bipartitions of the leaves the candidate trees $\cT$ can induce. 
A \textit{bipartition} of a set $S$ is a partition of $S = A \cup B$ into two nonempty, disjoint subsets. 
Every edge $e \in E(\cT)$ of our tree induces a bipartition of the leaves by considering the two connected components
of the graph $(V(\cT), E(\cT) \setminus \{e\})$. Moreover, Buneman's splits-equivalence theorem shows $\cT$
is completely characterized by its set $\cX(\cT)$ of bipartitions. 

Since each $g_i$ is a noisy estimate of $\cT^*$, we 
might expect that $\cX(\cT^*) \subseteq \cX_g := \bigcup_i \cX(g_i)$, suggesting
the restricted search space
%
\[\Big\{\cT \; : \; \cX(\cT) \subseteq \bigcup_{i=1}^n \cX(g_i)\Big\}.\]
%
\noindent More generally, given any set $\cX$ of bipartitions, we can solve
%
%
\begin{equation}
    \hat{T} = \argmax_{\cX(\cT) \subseteq \cX}  \sum_{q \in \cQ(\cT)} W(q).
    \label{eqn:astral_heuristic_optimization_problem}
\end{equation}
%
%
Using a clever dynamic-programming principle, ASTRAL efficiently solves this bipartition-constrained 
optimization problem. For polynomially-sized $\cX$, the optimization problem~\eqref{eqn:astral_heuristic_optimization_problem}
can be feasibly solved for trees with thousands of leaves. Moreover, when the trees $g_i$ are generated under the 
multispecies coalescent model \cite{rannala2020msc}, \cite{hill2020speciestreeestimationjoint}, its statistical properties 
are well-understood. In particular, it is asymptotically consistent whenever $\cX_g \subseteq \cX$. 

The paper~\cite{shekhar2018_how_many_genes_are_enough} showed for the exact version~\eqref{eqn:astral_quartet_support_optimization_problem}
that $n = \Theta(f^{-2} \log(k))$ independent gene trees $g_i$ are required to recover $\cT^*$ with high probability, where $f$ is the minimum
branch length in the tree. It has been empirically observed that the relaxed version~\eqref{eqn:astral_heuristic_optimization_problem} requires
significantly more data to obtain similar recovery rates~\cite{shekhar2018_how_many_genes_are_enough}. Thus there is a trade-off
between computational and sample efficiency.

As discussed in \cite{uricchio2016_bipartition_cover}, \cite{mcnulty2026improvedbipartitioncoverbound}, the main shortcoming of the relaxed problem
is that it is quite unlikely that $\cX(T) \subseteq \cX_g$, since the bipartitions induced by edges near the root are very 
difficult to recover. Several heuristics have been proposed to extend the set $\cX_g$ in an effort to improve its coverage 
properties and the empirical performance of the relaxed problem~\cite{mirarab2015astral}, \cite{zhang2018astral}, \cite{dibaeinia2021fastral}. 
Alternatively \cite{Zhang2022_weighted_ASTRAL} proposes a weighted version of~\eqref{eqn:astral_quartet_support_optimization_problem}
which incorporates quartet-support information from the gene-tree estimation stage. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ASTRAL-Based Inference for Canby Model
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PCH-ASTRAL: Phylogenetic Inference in the Canby Model}

For data generated under the Canby model, quartet support can be defined directly from the
shared non-homoplastic states of the observed languages. Let $q = uv||xy$ denote the unrooted
quartet topology that groups $u$ with $v$ and $x$ with $y$.

\begin{center}
\begin{tikzpicture}[scale=0.9, every node/.style={font=\small}]
    \coordinate (a) at (-1, 0);
    \coordinate (b) at (1, 0);
    \coordinate (u) at (-2.5, 0.9);
    \coordinate (v) at (-2.5, -0.9);
    \coordinate (x) at (2.5, 0.9);
    \coordinate (y) at (2.5, -0.9);

    \draw (u) -- (a) -- (b) -- (x);
    \draw (v) -- (a);
    \draw (y) -- (b);

    \node[left] at (u) {$u$};
    \node[left] at (v) {$v$};
    \node[right] at (x) {$x$};
    \node[right] at (y) {$y$};
\end{tikzpicture}
\end{center}

\noindent Given a character $c$, define the support $M^{(c)}_q$ of the quartet $q = uv||xy$ via
%
\begin{equation*}
    M^{(c)}_{q}:= \Big\{(\alpha, \beta) \in \Sigma_N \times \Sigma_N \; : \; \alpha \neq \beta, \quad \alpha \in u \cap v \cap \bar{x} \cap \bar{y}, \quad \beta \in \bar{u} \cap \bar{v} \cap x \cap y\Big\}. 
\end{equation*}
%
Namely, $M_q$ consists of the pairs of non-homoplastic states that are seemingly ``split'' across the quartet $q$.
For the true quartet topology, this support is larger in expectation than the support assigned to
the two alternatives:
%
\begin{lemma}
    If $\cT_q^* = uv||xy$ is the quartet topology of the leaves $(u,v,x,y)$ in $\cT^*$ then
    \[\E M_{uv || xy} > \E M_{ux || vy} = \E M_{uy || vx},\]
    %
    \noindent so the correct topology has the highest expected support.
    \label{lem:correct_topo_max_expected_support}
\end{lemma}
%
\noindent For each quartet $q$, they define its (empirical) support via
%
\begin{equation}
    S_q := \frac{1}{n} \sum_{i=1}^n |M_q^{(c_i)}|.
    \label{eqn:quartet_support_averaged}
\end{equation}
%
\noindent For each $s = \{u,v,x,y\} \in {[k] \choose 4}$ they then let
%
\begin{equation}
    q_s := \argmax_{q \in A_s} S_q,
    \label{eqn:astral_based_canby_inference_original}
\end{equation}
%
\noindent by the quartet topology with maximal empirical support. By the law of large numbers, the correct 
topology becomes overwhelmingly likely for large $n$. They then run ASTRAL on the quartet set $\{q_s\}$
to generate an estimate $\hat{\cT}$ of $\cT^*$. The authors call this algorithm \textit{PCH-ASTRAL}, and 
additionally propose a weighted version in \cite{evans2026_canby_identifiability}, although we will
only focus on the unweighted version described above.

Although not directly mentioned in \cite{evans2026_canby_identifiability}, the proof of Lemma~\ref{lem:correct_topo_max_expected_support}
directly implies the following. Recall, a phylogenetic tree (with branch lengths) is said to be \textit{ultrametric} if every leaf is
the same distance from the root. When this holds, the supports for the two incorrect topologies are identically distributed. 

\begin{corollary}
    If $\cT^*$ is ultrametric, then $M_{ux|vy} \overset{d}{=}M_{uy|vx}.$
    \label{corr:wrong_quartet_supports_identically_distributed}
\end{corollary}


%%%%%%%%%%%%
% Summary of contributions
%%%%%%%%%%%%
\subsection{Summary of Contributions}

In Section~\ref{sec:basic_results}, we construct a coupling of the Canby model to a simple immigration-death
process which allows for many direct computatins and will be the basis of most of our rigorous analysis. 
Using this, we develop some basic properties of both the original Canby model and of this coupled process.


Afterwords, in Section~\ref{sec:bipartition_covers_canby}, we show how the search-space for PCH-ASTRAL can be 
significantly restricted while maintaining statistical consistency. This vastly improves the run-time of the underlying
ASTRAL procedure, enabling phylogenetic analysis of hundreds of different languages simultaneously. While the 
restriction reduces the algorithms sample efficiency, we further introduce two heuristics meant to mitigate this
degradation in performance.

Following this, in Section~\ref{sec:canby_alternative_astral_procedure}, we explore an alternative ASTRAL-based 
inference procedure, which we call \textit{PCH-ASTRAL(sc)}. This procedure keeps character-level quartet information 
separate by converting each informative character-quartet pair into a single topology vote, rather than averaging
 support magnitudes across characters. While likely less statistically efficient, we argue that PCH-ASTRAL(sc) 
 may be more robust in heterogeneous or misspecified settings due to the heavy-tailed nature of the quartet-support 
 distributions.

Lastly, in Section~\ref{sec:pch_astral_sample_complexity}, we develop rudimentary sample complexity bounds for the two ASTRAL-based procedures
when characters are sampled under the simplified immigration-death process $ID(\theta, \lambda, \mu)$. 
We show PCH-ASTRAL requires $O(L^{-2} \log(k) + L^{-1} \log^2(k))$ characters to recover $\cT^*$ with high probability, where
%
    \[L := \min_{s \in {[k] \choose 4}} \min_{t \in A_s \setminus \{\cT_s^*\}} \E \left[|M_{\cT_s^*}| - |M_t|\right]\]
%
is the minimal margin the correct quartet topology has over incorrect ones. Similarly, 
PCH-ASTRAL(sc) requires $O(\Delta^{-2} \log(k))$ \textit{informative} characters, where
%
    \[\Delta := \min_{s \in {[k] \choose 4}} \min_{t \in A_s \setminus \{\cT_s^*\}} \left[\P_s(\cT_s^*) - \P_s(t)\right], \quad \P_s(t) := \P\left(t = \argmax_{t' \in A_s} |M_{t'}| \; \Big\vert \; I_s\right).\]
%
again gives some notion of the margin. 


%%%%%%%%%%%%
% Notation
%%%%%%%%%%%%
\subsection{Notation}

Let $\cT^*$ denote the true species tree underlying our Canby model,
which has $k$ leaves (species/languages). Given any phylogenetic tree $\cT$, let $\cQ_T$ denote its set of 
\textit{quartets}, which are the subtrees induced by a collection of four leaves. Hence $|\cQ_\cT| = {k \choose 4}$.

We will always use $q = uv || xy$ to denote a quartet, which inherently comes with some (unrooted) topology, and $s = \{u,v,,x,y\}$ 
to denote its underlying set of four leaves, which may not have a topology. Given a leaf set $s = \{u, v, x, y\}$, we will use 
%
\[A_s := \{uv||xy, ux||vy, uy||vx\}\] 
%
to denote the set of three different (unrooted) quartets on the same leaf set. We will use $\cT_s^*$ to denote the quartet $q \in A_s$
which appears in the true phylogenetic tree $\cT^*$. 


Unless stated otherwise, assume the observed characters are independent and identically distributed, and assume any
states $\alpha, \beta$ referenced are non-homoplastic. 

For ease of notation, when it is clear from contex we will denote the macro-state $c(v)$ at a given point $v \in \cT^*$ using upper-case letters. 
In particular, we will use $I := c(i), J := c(j),$ and $K := c(k)$ down below.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Basic Results}
\label{sec:basic_results}

%%%%%%%%%%%%%%%%%%%%%
% Coupling to ID Process
%%%%%%%%%%%%%%%%%%%%%
\subsection{Coupling to Immigration-Death Processes}

The constraint $\mu_1 = 0$ makes the Canby model hard to analyze, because it adds a certain
dependence structure between the different states of a specific character. If we drop this condition, we 
get a classical immigration-death (ID) process, which is quite tractable to analyze. This approximation
is reasonable in the parameter regime where the number of characters in a given branch is rarely one. 

Our key tool will be a coupling of the Canby model to a standard immigration-death process (ID)
 run along the branches of our species tree $\cT^*$. It relies on two simple observations 
 about how the set of nonhomoplastic states evolves over time:

\begin{itemize}
    \item Births always occur at rate at least $\lambda_N := \lambda \nu(\Sigma_N)$ and at most $\lambda$.

    \item Deaths always occur at rate at most $\mu$ per individual.
\end{itemize}
%
\noindent In particular, we can couple an ID process $(Y_s)_{s \in \cT^*}$ to our Canby $(X_s)_{s \in \cT^*}$ so that:
%
\[ Y_s \cap \Sigma_N \subseteq X_s \cap \Sigma_N, \quad \forall s \in \cT^*\]
%
\noindent In this way, we only reduce the number of nonhomoplastic states we observe in the leaves, which heuristically makes inference more challenging. 

To produce this coupling, use a single $\tau \sim \mathrm{\Exp}(\lambda)$ clock for each branch of $\cT^*$ to determine possible births. If $\tau$ rings at location $s \in \cT^*$, sample a new state $\alpha \notin X_{s^-}$ for the Canby model as usual, namely according to the conditional distribution:
%
\[\frac{\nu(\cdot \cap (\Sigma \setminus X_{s^-}))}{\nu( \Sigma \setminus X_{s^-})}\]
%
\noindent Hence we let $X_s = X_{s^-} \cup \{\alpha\}$. We update our ID process $Y_s$ as follows: if $\alpha \notin \Sigma_N$, then let $Y_s = Y_{s^-}$. Namely, ignore the homoplastic states. Otherwise:
%
\[Y_s = \begin{cases}
    Y_{s^-} \cup \{\alpha\} & \mathrm{w.p.}  \quad \nu(\Sigma \setminus X_{s^-})\\
    Y_{s^-} & \mathrm{w.p.} \quad 1-\nu(\Sigma \setminus X_{s^-}) \\
\end{cases}\]
%
\noindent This ensures that the birth rate in $(Y_s)$ along a given branch is always $\lambda_N$. 

Similarly, when a new state $\alpha$ is born in $X_s$, assign it an independent $\tau_\alpha \sim \mathrm{Exp(\mu)}$ clock. If it rings at location $s \in \cT^*$, let:
%
\[X_s = \begin{cases}
    X_{s^-} \setminus \{\alpha\} & X_{s-} \neq \{\alpha\}\\
    \{\alpha\} & \mathrm{otherwise}
\end{cases}\]
%
\noindent Then let $Y_s = Y_{s^-} \setminus \{\alpha\}$. Namely, $(X_s)$ simply ignores the deaths causing the state set to become empty, while $(Y_s)$ does not. 

Heuristically, this coupling well-approximates the non-homoplastic states of the Canby model in the high-polymorphism regime $\lambda \gg \mu$ where it is rare for there to be just a single 
state, and when there is little homoplasy so the renormalization of $\nu$ does not drammatically affect the birth-rate of non-homoplastic states over time. 

In Appendix~\ref{app:canby_id_proofs}, we collect some properties of these immigration-death processes which will be useful for
analyzing the Canby model. In particular, the following corollary of Lemma~\ref{lem:quartet_supports_under_ID_process}
shows bounding the quartet-supports reduces to understanding
some Poisson random variables.

\begin{corollary}
    Under the coupled $ID(\theta, \mu, \lambda)$ process, the quartet-supports equal
    %
        \[|M_{uv||xy}| = N_{uv} \cdot N_{xy}, \quad |M_{ux||vy}| = N_{ux} \cdot N_{vy}, \quad |M_{uy||vx}| = N_{uy} \cdot N_{vx}, \]
    %
    \noindent for $N_{uv}, N_{xy}$ independent Poisson random variables.
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%
% properties of Canby Model
%%%%%%%%%%%%%%%%%%%%%
\subsection{Properties of Canby Model}

In this section, we derive some basic properties of the Canby model. The derivations often rely on the 
coupling we introduced in the previous section. 

Given a path $[0, \ell]$ from the root of $\cT^*$ to some leaf, let $S_t = |c(t)|$ be the number 
of states in the macro-state. This is a simple birth-death process with birth rates 
$\lambda_i \equiv \lambda$ and death rates $\mu_i = i\mu \cdot 1\{i > 1\}$. Hence

% stationary distribution
\begin{lemma}
    The stationary distribution of $S_t$ is $ \pi = (Z | Z > 0)$ for $Z \sim \mathrm{Poisson}(\lambda/\mu)$. 
    %
    \[\pi_i = \frac{(\lambda/\mu)^i}{i!(e^{\lambda/\mu}-1)}, \quad \forall i \geq 1.\]
    %
    \label{lem:stationary_dist_state_cardinality_process}
\end{lemma}

\noindent The ergodic theorem tells us as $T \to \infty$ the fraction of the time $\{S_t\}_{t \in [0,T]}$ spends in state 1 
converges almost surely to $\pi_1 = \frac{\lambda / \mu}{e^{\lambda / \mu}-1}$. Heuristically, over long-branches our effective death-rate then becomes:
%
\[\mu_{\text{eff}} = \mu \left(1-\pi_1\right) = \mu \cdot \frac{e^{\lambda / \mu}-\lambda/\mu-1}{e^{\lambda / \mu}-1}\]
%
\noindent If $\lambda \ll \mu$, then $\mu_{\text{eff}} \approx \frac{\lambda}{2} $ which 
is significantly less than $\mu$ in this low-polymorphism regime.
If $\lambda \gg \mu$, then $\pi_1$ is exponentially small in $m$, so the 
process is rarely in state $1$. Hence it becomes well-approximated by our 
ID process coupling. 


For the following, we will also need two-sided bounds on the probability a given state $\alpha$ 
dies along a specific branch, which the next lemma provides.

\begin{lemma}[Survival and Death Bounds]

    Let $p_{n,T}$ be the probability that, starting with $n$ states, a given state $\alpha$ dies along a branch of length $T$ in the Canby model. Then
    %
    \[\frac{1}{2} (1-e^{-\lambda T/2}) (1-e^{-\mu T}) =:\xi_{T}\leq p_{n,T} \leq 1-\eta_{T} := 1-e^{-\mu T}\]
    %
    \label{lem:death_rate_prob_bounds}
\end{lemma}

Thus $\xi_T, \eta_T$ give lower bounds on the death and survival rates respectively.
 We will often be interested in applying these bounds to a specific path
 in the tree. Given two vertices $i, j$ in the tree denote by $\ell_{ij}$ the length 
 of the path $[i,j]$ in the tree and denote $\xi_{ij} := \xi_{\ell_{ij}}, 
 \eta_{ij} := \eta_{\ell_{ij}}$ our corresponding lower bounds on death and survival. 
 
 Next, we aim to get some control over the tails of our supports $|M_q|$. For $q = uv ||xy$, it is easy
 to see $M_q \subseteq (U \cap V) \times (X \cap Y)$. By Lemma~\ref{lem:stationary_dist_state_cardinality_process},
 we roughly see then that $|M_q|$ is the product of two Poisson random variables, whose correlation depends on 
 how closely related $u,v$ are with $x,y$. This suggests tails $\P(|M_q| > t) \leq \exp(-O(\sqrt{t} \log t))$, namely
slightly lighter tails than Weibull($1/2$) essentially. The paper \cite{kuchibhotla2022weibull} develops many useful concentration inequalities
for such random variables, which we rely on in Lemma~\ref{lem:canby_concentration_of_scores} to prove an analogous statement
for the ID process coupling. 

 Later on, we will need to develop some control over the expected margin 
 between the species tree topology and the other topologies:
    %
    \[\Delta^{(c)}_q := \E D_t, \quad D_t := |M^{(c)}_{T^*_q}| - |M_{t}^{(c)}|, \quad t \in \cT_q \setminus \{T^*_q\}\]
    %
\noindent which quantifies the gap in Lemma \ref{lem:canby_expected_margin}. Let $\rho_N = \E |c(r) \cap \Sigma_N|$ be the 
expected number of non-homoplastic states present at the root. Note for a given (rooted) quartet, there
are two possible topologies, as shown in Figure~\ref{fig:canby_quartet_topologies}. As denoted in the figure, let
$i$ denote the root, $j$ the most recent common ancestor of $u,v$, and $k$ the most recent common ancestor of $x,y$. 

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[
          level distance=12mm,
          level 1/.style={sibling distance=24mm},
          level 2/.style={sibling distance=14mm},
          every node/.style={font=\small},
          edge from parent/.style={draw,-},
        ]
        \node (i) {$i$}
          child { node (j) {$j$}
            child { node (u) {$u$} }
            child { node (v) {$v$} }
          }
          child { node (k) {$k$}
            child { node (x) {$x$} }
            child { node (y) {$y$} }
          };
        \end{tikzpicture}
        \caption{Balanced quartet.}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[
          level distance=12mm,
          level 1/.style={sibling distance=24mm},
          level 2/.style={sibling distance=14mm},
          every node/.style={font=\small},
          edge from parent/.style={draw,-},
        ]
        \node (i) {$i$}
          child { node (u) {$u$}}
          child { node (j) {$j$}
            child { node (v) {$v$} }
            child { node (k) {$k$}
                child { node (x) {$x$} }
                child { node (y) {$y$} }
            }
          };
        \end{tikzpicture}
        \caption{Caterpillar quartet.}
    \end{subfigure}
    \caption{The two rooted quartet topologies with split $uv||xy$. }
    \label{fig:canby_quartet_topologies}
\end{figure}


\begin{lemma}[Expected Margin Lower Bound] 
    
    If $q$ induces a balanced tree in $\cT^*$: 
    \[\Delta^{(c)}_q \geq \left(\frac{\lambda_N}{\mu} \right)^2(1-\eta_{ij})(1-\eta_{ik}) \eta_{uv} \eta_{xy}   \]

\noindent If $q$ induces a caterpillar tree and $r$ denotes the root of $\cT^*$, then
%
\[\Delta^{(c)}_q \geq  \frac{\lambda_N}{\mu} \left[\rho_N \eta_{ri} + \frac{\lambda_N}{\mu}(1-\eta_{ri})\right]  (1-\eta_{jk})\xi_{jk} \eta_{uv}\eta_{xy} \]

    \label{lem:canby_expected_margin}
\end{lemma}

\noindent More granular bounds can be found in the proof of Lemma~\ref{lem:canby_expected_margin} in the Appendix. 
The above highlights exactly the quartets that are hard to recover: 
those where $u,v$ and $x,y$ are far from their respective most 
recent common ancestors $j,k,$ and where these common ancestors 
$j,k$ are close to their least common ancestor $i$. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bipartition Covers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Restricted Search-Space}
\label{sec:bipartition_covers_canby}

\noindent As we discussed in the introduction, the exact version of ASTRAL~\eqref{eqn:astral_quartet_support_optimization_problem}
becomes intractable to solve even for trees with a modest number of leaves. Instead, we must construct a 
set $\cX$ of bipartitions which will limit the search space for the true tree $\cT$, which is the content of this section. 

To this effect, one useful class of sets when it comes to building $\cX$ are the leaf sets:
%
\[L_\alpha := \{\text{leaves }\ell : \alpha \in c(\ell)\}\]
%
\noindent Say a subset $L \subseteq S$ of leaves is \textit{contained} within a given bipartition $\phi := (A|B)$ if:
%
\[L_\alpha \subseteq A \qquad \text{or} \qquad L_\alpha \subseteq B\]
%
It is easy to see for nonhomoplastic states $\alpha \in \Sigma_N $, if $\alpha$ is born
 on the edge $e$ and $\phi_e = (A_e | S-A_e)$ is the bipartition this edge 
 generates in $\cT^*$ then $L_\alpha$ is contained in $\phi_e$. Moreover, if no
deaths of $\alpha$ occur after its born then in fact $L_\alpha \in \{A_e, S-A_e\}$. Hence 
one choice for the candidate bipartitions is
%
\[\cX_1 := \{(L^{(c)}_\alpha|S-L_\alpha) : c \in C, \alpha \in \Sigma^{(c)}\}\]
%
\noindent At stationarity, we typically have around $\lambda \mu^{-1}$ states in each leaf,
so roughly we can see $|\cX_1| \lesssim k |C| \lambda \mu^{-1}$ grows only linearly in $k$, so its often 
much smaller than the total $2^k$ number of possible bipartitions. It is easy to see under mild 
assumptions this alone is sufficient for ensuring statistical consistency, 
since with positive probability a given non-homoplastic $\alpha$ is born 
along an edge $e$ and does not die at all. More granullary

% bipartition cover bound
\begin{theorem}[Bipartition Cover Bounds for Canby With No Borrowing]
Let $\cB$ be the set of bipartitions of a phylogenetic tree $\cT$ with $k$ leaves and minimum/maximum branch lengths $T_{min}, T_{max}$ respectively. Suppose further we have a Canby model $\left\{\Sigma^{(c)}, \lambda^{(c)}, \mu^{(c)}, \rho^{(c)}\right\}_{c \in C}$ on $\cT$ with no borrowing and whose rates follow an immigration-death process. Let:
%
\[C_N := \{c \in C \; : \; \nu^{(c)}(\Sigma_N^{(c)}) > 0\}\]
%
\noindent be the set of characters with positive chance of non-homoplasy. Furthermore, assume we have the following assumptions on the parameters of the Canby model:

\begin{enumerate}
    
    \item[(i)] \textit{Nonvanishing Birth Rates:} Suppose that $\lambda^{(c)} \nu^{(c)}(\Sigma_N^{(c)}) \geq b > 0$ for all $c \in C_N$.

    \item[(ii)] \textit{Bounded death rates}: Suppose $\mu^{(c)} \leq D < \infty$ for all $c \in C_N$. 
\end{enumerate}
%
\noindent Then we have exponential convergence of the coverage probability:
%
\[\P(\cB \nsubseteq \cX_1)  \leq\sum_{\ell=2}^{k-2} q_\ell^{|C_N|} \leq (k-3)q_{k-2}^{|C_N|}\qquad q_\ell := \exp\left(-bT_{min} \exp\left[-2\ell DT_{max}\right]\right) \]
%
\noindent In particular, we have asymptotic consistency when $|C_N| \to \infty$:
%
\[\lim_{|C_N| \to \infty} \P(\cB \subseteq \cX_1) = 1,\]
%
\label{thm:canby_bipartition_cover_bounds_no_borrowing}
\end{theorem}
%
In particular, this implies if we run the ASTRAL as proposed in \cite{evans2026_canby_identifiability},
restricted to the bipartition set $\cX_1$, this is a statistically consistent estimate 
of the true underlying phylogenetic tree $T$, and that we require $\exp(\Theta(k))$ characters to
achieve a bipartition cover with high probability. This is essentially the same rate
bipartition-covers achieve under the standard multispecies-coalescent model to which
ASTRAL is typically applied \cite{uricchio2016_bipartition_cover}, \cite{mcnulty2026improvedbipartitioncoverbound}.

This result should be viewed as a baseline consistency guarantee rather 
than a practical sample-complexity result. It is clear that $\cX_1$ is 
bad at generating large bipartitions.
 Namely, if $\phi_e = (A|B)$ the probability a 
  character born in edge $e$ has no deaths is exponentially
   small in $|A|$. To rectify this, we propose a few heuristics for expanding the size
    of $\cX_1$ so it has better coverage probabilities. A natural solution
     is to allow multiple states $\alpha$ to contribute:
%
\[\cX_m := \left\{\bigcup_{i=1}^r L_{\alpha_i}^{(c)}  \;: \; c \in C, \alpha_i \in \Sigma^{(c)}, r \leq m\right\}\]
%
\noindent It is easy to see that $\cX_n \subseteq \cX_m$ for $n \leq m$, so in
 particular $\cX_1 \subseteq \cX_m$. Hence asymptotic consistency using $\cX_m$ follows 
 from that of $\cX_1$. To avoid combining leaf sets from 
 completely different regions in the tree, we can also consider some sort of compatibility condition like
 $L_{\alpha_i}^{(c)} \cap L_{\alpha_j}^{(c)} \neq \emptyset$.

Alternatively, we can look for sets $A$ so that $L_\alpha$ takes up ``most''
 of $A$. Formally, defining some threshold $\tau$ we let
%
\[\cX_\tau := \left\{(A|S-A) \; : \; \exists \alpha \; \left(\frac{|A \setminus L_\alpha|}{|A|} < \tau\right)\right\}\]
%
\noindent Since $|A| < k$, this allows at most $\tau k$ additions. 
Hence if we choose $\tau = r/k$ for $r < k$, we limit ourselves to at most $r$ 
additions. 



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternative ASTRAL procedure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternative ASTRAL Procedure}
\label{sec:canby_alternative_astral_procedure}

Another inference method, closer in spirit to the original ASTRAL framework, is to let each
character independently vote on the topology of each quartet. These character-level quartet
votes then play the role of the quartet information induced by individual gene trees in the
traditional MSC setting.

More formally, for each character $c_i$ and each leaf set
$s=\{u,v,x,y\}$, define
\[
    \tau_s^{(c_i)}
    :=
    \arg\max_{q\in\mathcal A_s} |M_q^{(c_i)}|, \quad I^{(c_i)}_s := \left\{\max_{q\in\mathcal A_s} |M_q^{(c_i)}| > 0\right\},
\]
\noindent with ties broken uniformly. We then solve~\eqref{eqn:astral_quartet_support_optimization_problem} with 
quartet supports
%
    \[W(q) = \sum_{i=1}^n 1\{\tau_s^{(c_i)} = q\} \cdot 1\{I_s^{(c_i)}\}.\]
%
The indicator $1\{I_s^{(c_i)}\}$ essentially just tells us to ignore characters which provide no information
about the given quartet. We call this procedure \textit{PCH-ASTRAL(sc)} for ``separate characters''. 
This inference scheme is only really reasonable if for 
each quartet $q \in \cQ(\cT^*)$, the correct topology is the most likely one. Namely

\begin{equation}
    \P(\tau_s^{(c)} = \cT_s^*) > \P(\tau_s^{(c)} = t) \qquad t \neq T_s^*.
    \label{eqn:anomaly_free_property}
\end{equation}

\noindent If property~\eqref{eqn:anomaly_free_property} holds for every quartet $q \in \cQ(\cT^*)$,
we say the Canby model is in the \textit{anomaly-free zone}. When this occurs, the above 
ASTRAL-based procedure is asymptotically consistent and has sample-complexity

\begin{theorem}[Sample Complexity PCH-ASTRAL(sc)]
    Suppose we have a collection of $n$ iid characters $\{c_i\}$ in the anomaly-free zone. If
        \[n > \frac{9}{p_{min}\Delta^2} \log\left(\frac{4\binom{k}{4}}{\varepsilon}\right),\]
        %
        \noindent where
        %
        \[\Delta := \min_{s \in {[k] \choose 4}} \min_{t \neq \cT_s^*} \left[\P(\tau_s^{(c)} = \cT_s^* \mid I_s) - \P(\tau_s^{(c)} = t \mid I_s) \right], \quad p_{min} := \min_{s \in {[k] \choose 4}} \P(I_s),\]
        %
    \noindent then the above ASTRAL-based procedure recovers $\cT^*$ with probability at least $1-\epsilon$. 
    \label{thm:sample_complexity_separate_characters}
\end{theorem}

\noindent The proof is analogous to the proof of Theorem 2.1 in~\cite{shekhar2018_how_many_genes_are_enough}, which union bounds over quartets and
utilizes simple concentration bounds for the multinomial distribution. In Appendix~\ref{app:anomaly-free_zone}, we briefly discuss the
anomaly-free zone. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sample Complexity
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample-Complexity Bounds}
\label{sec:pch_astral_sample_complexity}

% ave across characters
\subsection{PCH-ASTRAL Sample-Complexity Bounds}

In this section, we aim to understand the number of independent characters required by PCH-ASTRAL
to recover the true species tree $\cT^*$ with high probability. To attempt to do so, we prove a sample complexity
result for the PCH-ASTRAL procedure when run on samples generated from an immigration-death process
$ID(\theta, \mu, \lambda)$ run on the true species tree $\cT^*$. It is a simple consequence of our concentration
bound~\ref{lem:canby_concentration_of_scores}.

\begin{lemma}[Sample Complexity PCH-ASTRAL under ID Process]
    Suppose we have a collection $\{c_i\}_{i=1}^n$ of iid characters whose multi-states are distributed according to the 
    $ID(\theta, \mu, \lambda)$ process we discussed above ran on the tree $\cT^*$. Let $\zeta = \theta \vee \frac{\lambda}{\mu}$ and
    %
    \[\kappa := \max_{s\in {[k]\choose 4}} \max_{uv\|xy\in A_s}\left(1+\zeta\eta_{uv}\right)\left(1+\zeta\eta_{xy}\right),\]
    %
    \noindent and define our worst-case margin
        \[
            L:=\min_{s\in {[k]\choose 4}}\min_{t\in A_s\setminus\{\cT_s^*\}}\E\left[|M_{\cT_s^*}|-|M_t|\right].
        \]
    \noindent Then if we have at least:
    %
    \[n \gtrsim \max\left\{\frac{\kappa^2}{L^2} t_\delta, \; \frac{\kappa}{L} t_\delta^2 \right\}, \quad t_\delta := \log\left(\frac{4{k \choose 4}}{\delta} \right), \]
    %
    \noindent characters, we recover the true species tree with probability at least $1-\delta$. 
    \label{lem:sample_complexity_pch_astral}
\end{lemma}

\noindent As discussed in the proof, the parameter $\kappa$ in a sense measures how heavy-tailed the support sizes $|M_{uv||xy}|$ are. 
It has the trivial upper bound $\kappa \leq (1 + \zeta)^2$, which is a function merely of the expected number of
states in a leaf of the $ID(\theta, \lambda, \mu)$ process. Hence the worst-case margin $L$ is the main parameter
of interest. Analogous to Lemma~\ref{lem:canby_expected_margin}, under $ID(\theta, \mu, \lambda)$ we can lower bound $L$ 
purely in terms of the underlying model parameters. 


The worst-case margin $L$ certainly depends on the topology of $\cT^*$, although we can roughly analyze
Lemma~\ref{lem:canby_expected_margin} to generate a heuristic topology-free bound. If $f$ is the minimal branch
length in $\cT^*$ and
%
    \[d := \max_{uv||xy \in \cQ(\cT^*)} \ell_{uv} \vee \ell_{xy},\]
%
\noindent is the maximal distance between any two leaves on the same side of a quartet, 
then the margin bounds in Lemma~\ref{lem:canby_expected_margin} roughly become
%
    \[\Delta_q \geq \left(\frac{\lambda}{\mu}\right)^2 (1-e^{-\mu f})^2 (e^{-\mu d})^2 \gtrsim \lambda^2 f^2 e^{-2\mu d}, \]
%
\noindent when the induced subtree of $q$ in $\cT^*$ is balanced and similarly
%
        \[\Delta_q \gtrsim \lambda^2 \mu f^3 e^{-2\mu d} \left(\rho_N \wedge \frac{\lambda}{\mu}\right), \]
%
\noindent when the induced subtree is a caterpillar. Thus the worst-case margins are dictated by its caterpillar trees,
and for a tree of a fixed height the bound scales roughly like
%
    \[n \gtrsim \max\left\{f^{-6} \log(k), \; f^{-3} \log^2(k)\right\},\]
%
\noindent significantly worse than the $\Theta(f^{-2} \log(k))$ of ASTRAL under the multi-species coalescent model,
as discussed in \cite{shekhar2018_how_many_genes_are_enough}. Of course, our bound is naturally quite loose, and future 
improvements may significantly improve the rate.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Discussion and Concluding Remarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Concluding Remarks}

In this chapter, we developed several potential candidates for bipartition covers which 
substantially reduce the search space of the original PCH-ASTRAL inference procedure. 
This modification plausibly expands the scope of PCH-ASTRAL from a few dozen languages 
to hundreds of languages. Although restricting the search space reduces sample efficiency, 
we proposed several heuristics designed to mitigate this loss in performance.

Beyond this, we derived preliminary sample-complexity bounds under the immigration-death 
process model $ID(\theta,\lambda,\mu)$, both for the original PCH-ASTRAL procedure and for 
the robustified procedure PCH-ASTRAL(sc). In both cases, the rates depend explicitly on a 
quartet-level margin measuring the advantage of the correct quartet topology over the 
incorrect ones. We further gave partial bounds relating these margins to the underlying 
model parameters.

There are several natural directions for future work. First, one would like to extend the
 sample-complexity results from the simplified $ID(\theta,\lambda,\mu)$ process to the 
 full Canby model. The main obstacle is concentration. The immigration-death coupling 
 provides a convenient lower bound on non-homoplastic state counts, which is enough 
 for expected-margin estimates, but the concentration argument also requires an 
 upper bound on macro-state sizes at the leaves. Establishing such an upper bound 
 for the full Canby process would therefore be an important first step toward extending
  Lemma~\ref{lem:sample_complexity_pch_astral} beyond the ID approximation.

Second, since the true tree $\cT^*$ is unknown at inference time, it would be useful 
to develop topology-free sample-complexity bounds depending only on simple geometric
 parameters of $\cT^*$, such as the minimum branch length, analogous to the results of 
 \cite{shekhar2018_how_many_genes_are_enough}. Under the ID approximation, the Poisson 
 means $m_{uv}=\E N_{uv}$ are explicit functions of the model parameters, so more careful
bookkeeping in Lemma~\ref{lem:sample_complexity_pch_astral} may lead to relatively explicit 
bounds. 

Third, the anomaly-free condition for PCH-ASTRAL(sc) remains somewhat opaque. Under the ID 
approximation, the condition reduces to a comparison of independent product-Poisson random 
variables, and is therefore numerically checkable from the quartet-level means. However, 
we do not currently have simple geometric sufficient conditions ensuring that the anomaly-free
 property holds. 

Finally, several refinements of the present analysis should be possible. The margin bounds 
used here depend on coarse survival and death probability estimates from 
Lemma~\ref{lem:death_rate_prob_bounds}; sharper estimates, for example using Laplace transforms
 of hitting times, may yield more granular constants. More extensive empirical simulations 
 would also be valuable, both to compare PCH-ASTRAL and PCH-ASTRAL(sc) under various 
 model regimes and to assess how well the theoretical margins predict finite-sample behavior.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 
\begin{subappendices}

\section{Immigration-Death Process Proofs}
\label{app:canby_id_proofs}

Let $\mathrm{ID}(\theta, \mu, \lambda)$ denote an immigration-death process $Y_t$ with $|Y_0| \sim \mathrm{Poisson}(\theta)$,
 birth rate $\lambda$, and death rate $\mu$ per individual. We start by understanding 
 the population size in such a process. 

%%%%%%%%%%
% Population Sizes ID Processes
%%%%%%%%%%

\begin{lemma}[Population Sizes in ID Processes]
    If $(Y_t)_{t \geq 0} \sim \mathrm{ID}(\theta, \mu, \lambda)$ then: 
    %
    \[|Y_t| \sim \mathrm{Poisson}(m_\theta(t)), \quad m_\theta(t) = \theta e^{-\mu t} + \frac{\lambda}{\mu}(1-e^{-\mu t}).\]
    %
    \noindent In particular, its stationary distribution is $\mathrm{Poisson}(\lambda/\mu)$. Moreover for $s < t$:
    %
    \[\E|Y_t \cap Y_s| = m_\theta(s) \cdot  e^{-\mu (t-s)}\]
    %
    \label{lem:pop_sizes_in_ID_processes}
\end{lemma}

\begin{proof}
    Kolmogorov's forward equations tell us:
    %
    \[\frac{d}{dt}p_k = \lambda p_{k-1} + \mu (k+1) p_{k+1} - (k\mu + \lambda) p_{k}, \quad p_k(0) = \frac{e^{-\theta} \theta^k}{k!}\]
    %
\noindent It is easy to see the Poisson probabilities $p_k(t) =  \frac{e^{-m(t)}m(t)^k}{k!}$ satisfy these equations, and hence $X(t) \sim \mathrm{Poisson}(m_\theta(t))$. The covariance formula follows from the fact:
\[|Y_t| = \mathrm{Binomial}(X_s, e^{-\mu (t-s)}) + |Z|, \quad |Z| \sim \mathrm{Poisson}(m_0(t-s))\]
\noindent where these two terms are independent. The formula follows from the law of total covariance. 
\end{proof}

\noindent It will be important to know how much time the Canby model spends with just a single state. For this purpose, it will be important to know how long it takes to go down from $n$ states to one. Since the dynamics of the ID process and Canby model essentially agree when there are more than one state, this reduces to studying the hitting times of our ID process. 


%%%%%%%%%%
% Hitting Times ID Process
%%%%%%%%%%
\begin{lemma}[Hitting Times in ID Process] Let:
%
\[\tau_n := \inf\{t > 0 : |Y_t| = 1\}\]
%
\noindent conditional on $|Y_0| = n$. Then: 
    %
    \[\E \tau_2 = \frac{\mu}{\lambda^2}(e^{\lambda / \mu} - 1 - \lambda/\mu)\]
    %
\noindent Moreover, its Laplace transform satisfies:
%
\[\frac{2\mu}{\lambda + 2\mu + s}\leq \E e^{-s\tau_2} \leq \frac{\lambda + 2\mu}{\lambda + 2\mu + s}\]
%
\end{lemma}

\begin{proof}
    Let $w_n := \E \tau_n$. Note by conditioning on the time of the first event we see:
    %
    \[w_n = \frac{1}{\lambda + n\mu} \left[1 + \lambda w_{n+1} + n\mu w_{n-1}\right], \quad w_1 = 0\]
    %
    \noindent 

    Next control the Laplace transform $\phi_n(s) := \E e^{-s \tau_n}$. We will rely on the classical relation between the Laplace transform of birth-death processes and continued fractions. Conditioning on the first event time (either a birth or a death) we see:
    %
    \[\phi_n = \frac{\lambda}{\lambda + n \mu} \cdot \phi_{n+1} + \frac{n\mu}{\lambda + n \mu} \cdot \phi_{n-1}, \qquad \phi_1 \equiv 1\]
    %
    \noindent Noting that $\phi_3 \in [0,1]$ gives the desired bound on $\phi_2$. More generally, this gives an upwards recurrence for the ratios $r_n := \frac{\phi_n}{\phi_{n-1}}$ where $r_2 := \phi_2$:
    %
    \[r_n = \frac{n\mu}{\lambda + n\mu + s - \lambda r_{n+1}} \]
    %
    \noindent Iterating gives the continued fraction expansion:
    %
    \[r_{n} = \frac{n\mu}{\lambda + n\mu + s - \frac{\lambda(n+1)\mu}{\lambda + (n+1)\mu + s - ...}} \]
    %
    \noindent We can truncate this at even or odd convergents to get lower and upper bounds respectively on $r_n$. From these, we can bound $\phi_n(s) = \prod_{k=2}^n r_k$. 
\end{proof}

\noindent Note that the above implies in the low-polymorphism regime $\lambda \ll \mu$, we get $\E \tau_2 = \frac{1}{2\mu} + O(\lambda/\mu^2)$. 

%%%%%%%%%%%%%%%
% Stationarity distribution Canby Cardinality Process
%%%%%%%%%%%%%%%
\begin{proof}[Proof of Lemma \ref{lem:stationary_dist_state_cardinality_process}]
    This process is reversible, so by detailed balanced:
    %
        \[\pi_{i+1}(i+1)\mu = \lambda \pi_i, \quad \forall i \geq 1\]
    %
    \noindent Recursively applying this shows $\pi_i \propto \frac{(\lambda/\mu)^i}{i!}$ as desired.
\end{proof}

%%%%%%%%%%%%%%%
% Canby Death Rate Bounds
%%%%%%%%%%%%%%%
\begin{proof}[Proof of Lemma \ref{lem:death_rate_prob_bounds}]
    Suppose along the given branch we spent a fraction $\tau \in [0,1]$ of the time with at least two states, whose distribution implicitly depends on $n,T$. Then if $\alpha$ is a state present at time zero:
    %
    \[\P(\alpha \text{ survives along branch} \mid \tau) = e^{-\mu T \cdot  \tau}\]
    %
    \noindent Hence the survival probability depends only on the Laplace transform $\phi$ of $\tau$:
    %
        \[\P(\alpha \text{ survives along branch}) = \phi(\mu T)\]
    %
    Trivially $\tau \leq 1$, which gives the upper bound. For the lower bound, note that $p_{n,T}$ is increasing in $n$ by a simple coupling argument. Hence WLOG we assume $n=1$. One way for $\alpha$ to die in this setting is for another state $\beta$ to be born at time $\gamma \sim \mathrm{Exp}(\lambda)$. The first of these two states dies by time $T$ with probability $1-e^{-2\mu(T-\gamma)}$ and the death is equally likely to be $\alpha$ or $\beta$. Hence:
    %
    \[p_{n,T} \geq \frac{1}{2} \cdot \int_0^T (1-e^{-2\mu (T-\gamma)}) \lambda e^{-\lambda \gamma} d\gamma = \frac{1}{2}\left[(1-e^{-\lambda T}) -  \frac{\lambda(e^{- \lambda T} - e^{-2\mu T})}{2\mu - \lambda}\right] \]
    %
    \noindent An even simpler bound is:
    %
    \[p_{n,T} \geq \frac{1}{2} (1-e^{-\lambda T/2}) (1-e^{-\mu T})\]
    %
    \noindent which is the probability we have a birth of some state $\beta$ by time $T/2$, a death of either $\alpha, \beta$ in the remaining time $T/2$, and that this death is $\alpha$. 
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Quartet Supports under ID(theta, lam, mu)
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent The following shows that our quartet supports $|M_{uv||xy}|$ take a particularly simple form
under the ID process coupling. 

\begin{lemma}[Quartet Supports under ID Process]
    Suppose $Y_t \sim ID(\theta, \lambda, \mu)$ on the underlying tree $\cT^*$. For any subset $\cL$
    of the leaves of $\cT^*$, for each $A \subseteq \cL$ define
    %
        \[N_A^\cL := \left|\left(\bigcap_{\ell \in A} \{\alpha \in Y_\ell\}\right) \cap \left(\bigcap_{\ell \notin A} \{\alpha \notin Y_\ell\}\right) \right|\]
    %
    \noindent the number of states $\alpha$ which appear exactly in the set of leaves $A$, among 
    those in $\cL$. Then $\{N_A^{\cL}\}_{A \subseteq \cL}$ are mutually independent Poisson random 
    variables. In particular, for $s = \{u,v,x,y\}$ this implies
    %
        \[|M_{uv||xy}| = N^{(s)}_{uv} \cdot N^{(s)}_{xy},\]
    %
    \noindent is the product of two independent Poisson random variables.
    \label{lem:quartet_supports_under_ID_process}
\end{lemma}

\noindent This Lemma makes analyzing the expected margin
%
    \[\Delta_t := \E |M_{\cT^*_s}| - \E |M_t|, \quad t \in A_s \setminus \{\cT_s^*\},\]
%
\noindent particularly easy under the $ID(\theta, \lambda, \mu)$ process: we just must try to 
understand the means $m^{(s)}_{uv}$ of these Poisson random variables $N^{(s)}_{uv}$. Calculating these exactly 
can be challenging, but Lemma~\ref{lem:pop_sizes_in_ID_processes} gives a simple upper bound
%
    \[m_{uv}^{(s)} \leq \left(\theta \vee \frac{\lambda}{\mu}\right)\eta_{uv}.\]
%
\noindent The first term accounts for the number of states reaching the most recent common ancestor of $u,v$, and
the second term thins these to only those reaching both $u,v$. 

\begin{proof}[Proof of Lemma \ref{lem:quartet_supports_under_ID_process}]
    Fix a point $x \in \cT^*$. For any states $\alpha$ born at $x$, we see that
    $\alpha \in N_A^{\cL}$ for some $A \subseteq \cL$. For fixed tree topology $\cT^*$ and model
    parameters $\theta, \lambda, \mu$, the probability $p_x(\alpha \in N_A^{\cL})$
    it lands in a particular one of these sets depends only on the location $x$, and is 
    independent of the birth-process itself and any other states. 

    The number of states born along the tree form a Poisson point process
    with intensity
    %
        \[\nu(dx) := \theta \delta_r(dz) + \lambda dz,\]
    %
    where $r$ is the root and \(dz\) denotes length measure on the branches of \(\cT^*\).
    Kingman's marking theorem then implies that the marked process forms a Poisson
    point process on $\cT^* \times \{A : A \subseteq \cL\}$ with intensity $\nu(dz) p_x(dA)$. See, 
    for example, the chapter on marked Poisson processes in \cite{Kingman_poisson_processes}.
    The desired result then follows immediately.
\end{proof}

%%%%%%%%%%%%%%%
% M_{uv|xy} tail bound
%%%%%%%%%%%%%%%

\begin{lemma}[Tail Bounds for Quartet Supports]
    Let $\zeta = \theta \vee \frac{\lambda}{\mu}$. Under $ID(\theta, \mu, \lambda)$
    %
    \[\P(|M_{uv||xy}| > t) \leq 2\exp\left(- \zeta \min\{\eta_{uv}, \eta_{xy}\}  \cdot h\left(\zeta^{-1}\sqrt{\frac{t}{\eta_{uv} \eta_{xy}}}\right)\right)\]
    %
    \noindent where $h(t) := (t\log(t) - t + 1) 1\{t \geq 1\}.$ Moreover, $M_{uv||xy}$ is sub-Weibull$(1/2)$
    with Orcliz pseudo-norm associated to $\Phi_{\alpha}(x) = \exp(x^\alpha) - 1$ given by
    %
    \[||M_{uv||xy}||_{\Psi_{1/2}} \lesssim \left(1 + \zeta \cdot \eta_{uv} \right) \left(1 + \zeta \cdot \eta_{xy} \right).\]
    %
    \label{lem:canby_concentration_of_scores}
\end{lemma}

\begin{proof}[Proof of Lemma~\ref{lem:canby_concentration_of_scores} (quartet-support tail bound)]
    Note that:
    %
        \[M_{uv||xy} \subseteq \{(\alpha, \beta) : \alpha \in  U \cap V, \beta \in X \cap Y\} = (U \cap V) \times (X \cap Y)\]
    %
    \noindent Suppose that $w, z$ are the most recent common ancestors of $u,v$ and $x,y$ respectively. Lemma~\ref{lem:pop_sizes_in_ID_processes} 
    implies that $|Z|, |W| \leq_{st} \mathrm{Poisson}(\zeta)$. Then Poisson-thinning implies
    %
    \[|U \cap V| \leq_{st} \mathrm{Poisson}\left(m_{uv}\right), \quad  m_{uv} := \zeta \eta_{uv},\]
    %
    \noindent and analogously for $|X \cap Y|$. Chernoff's implies:
    %
    \[\P(\mathrm{Poisson}(m) \geq mt) \leq \exp(-m h(t)), \]
    %
    \noindent By union bound, for any $\gamma > 0$:
    %
    \[\P(|M_{uv||xy}| > t) \leq \P(|U \cap V| \geq \gamma\sqrt{t}) + \P(|X \cap Y| \geq \gamma^{-1}\sqrt{t}).\]
    %
    \noindent Setting $\gamma = \sqrt{m_{uv} /m_{xy}}$ weights the two terms according to their means, giving:
    %
    \[\leq \exp\left(-m_{uv} h \left(\sqrt{\frac{t}{m_{uv} m_{xy}}}\right)\right) + \exp\left(-m_{xy} h \left(\sqrt{\frac{t}{m_{uv} m_{xy}}}\right)\right)\]
    %
    \noindent from which the desired bound directly follows. 

    To upper bound the Orcliz norm, we rely on the H\"older-type inequality appearing as 
    equation~(3.5) in~\cite{kuchibhotla2022weibull}. This gives
    %
    \[|| M_q ||_{\Psi_{1/2}} \leq || \mathrm{Poisson}(m_{uv})||_{\Psi_1} \cdot || \mathrm{Poisson}(m_{xy})||_{\Psi_1}.\]
    %
    \noindent A simple calculation and the bound $\log(1+x) \geq \frac{x}{1+x}$ for $x \geq 0$ shows
    %
    \[||\mathrm{Poisson}(m) ||_{\Psi_1} = \frac{1}{\log(1 + \log(2) / m)} \leq 1 + \frac{m}{\log(2)}\]
    %
    \noindent from which the result follows.
\end{proof}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Anomaly-free zone
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Anomaly-Free Zone}
    \label{app:anomaly-free_zone}

    In this section, we briefly discuss the anomaly-free property~\eqref{eqn:anomaly_free_property}. 
    Lemma~\ref{lem:correct_topo_max_expected_support} shows $\E |M_t|$ is maximized over $t \in A_s$
    when $t = \cT_s^*$, but this does not imply $\cT^*$ maximizes
    %
        \[\P\left(t = \argmax_{t' \in A_s} |M_{t'}|\right), \] 
    %
    \noindent as it is possible $|M_{\cT_s^*}|$ is zero most of the time, but takes on large values
    rarely. To avoid this, we must show the supports $|M_{\cT_q^*}|$ are reasonably concentrated.
    We focus on the simpler $ID(\theta, \lambda, \mu)$ setting. 
    
    Lemma~\ref{lem:quartet_supports_under_ID_process} shows in this setting that the supports take the particular form
    %
        \[|M_{uv||xy}| = N_{uv}N_{xy}, \quad |M_{ux||vy}| = N_{ux} N_{vy}, \quad |M_{uy||vx}| = N_{uy} N_{vx},\]
    %
    \noindent where these six random variables $\{N_{uv}, ..., N_{vx}\}$ are mutually independent Poissons. Their means $m_{ab} := \E N_{ab}$ 
    are reasonably simple to compute. For example, we know from Lemma~\ref{lem:pop_sizes_in_ID_processes} that $|I| \sim \mathrm{Poisson}(m_\theta(\ell_{ri}))$.
    Hence if the true quartet topology is $q = uv||xy$, and it induces a balanced subtree then
    %
        \[m_{uv} = \eta_{uv} \left[m_0(\ell_{ij}) + m_\theta(\ell_{ri}) \eta_{ij}(1 - \eta_{ik} (\eta_{ky} + \eta_{kx} - \eta_{kx} \eta_{ky}))\right], \]
        \[m_{ux} = m_\theta(\ell_{ri})\eta_{ux} \left[(1-\eta_{jv}) \cdot (1-\eta_{ky})\right], \]
    %
    \noindent The main difficultly is that these means depend on the underlying parameters $(\cT^*, \lambda, \mu, \theta)$ 
    in complicated, inter-woven ways.

    \red{finish}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Additional Proofs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Additional Proofs}
\label{app:canby_additional_proofs}

%%%%%%%%%%%%%%%
% Bipartition Cover Bounds
%%%%%%%%%%%%%%%
\begin{proof}[Proof of Theorem \ref{thm:canby_bipartition_cover_bounds_no_borrowing} (Bipartition Cover Bound)]

Fix an edge $e \in E(\cT)$ with $k_e$ descendants and a given character $c$. For ease of notation, we drop all the $(c)$ superscripts for now. If $\phi_e$ is the corresponding bipartition, we will lower bound the probability that $\phi_e = (L_\alpha|S-L_\alpha)$ for some $\alpha \in \Sigma$.

First, note that as the birth-rate is constant the number $N_e$ of non-homoplastic states born along the edge $e$ satisfies:
%
\[N_e \sim \mathrm{Poisson}(\lambda T_e \cdot \nu(\Sigma_N)) \succeq_{st} Poisson(\eta) \qquad \eta := \lambda T_{min} \nu(\Sigma_N)\]
%
\noindent where $T_e$ is the length of the branch $e$ and $T_{min}$ the minimum branch length in the tree. 

For a given non-homoplastic $\alpha$ born along the edge $e$, we can lower bound the probability it has no deaths below $e$. By Lemma \ref{lem:death_rate_prob_bounds} the probability $\alpha$ survives along a branch of length $T$ is at least $\exp(-\mu T)$. There are $2k_e - 1 \leq 2k_e$ edges (including $e$) that $\alpha$ can die on. Each is at most length $T_{max}$, so we can lower bound the probability $\alpha$ has no deaths at all by:
%
\[\P(\alpha \text{ has no deaths} \; |\alpha \in \Sigma_N, \alpha \text{ born on e}) \geq \exp(- \tau_{k_e}) \qquad \tau_{k_e} := 2\mu k_e T_{max}\]
%
\noindent Using this result followed by $N_e \succeq_{st} \mathrm{Poisson}(\eta)$ we see:
%
\[\P(\exists\alpha \; \phi_e = (L_\alpha, S-L_\alpha)) \geq \E_{n \sim N_e} \left(1-(1-\exp(-\tau))^n\right) \geq 1-\exp(-\eta e^{-\tau})\]
%
 From here, we use the caterpillar bound from~\cite{mcnulty2026improvedbipartitioncoverbound} applied to our $|C|$ characters implies:
%
\[\P(\cB \subseteq \cX_1) \geq 1 - \sum_{\ell=2}^{k-2} \prod_{c \in C}\exp\left(-\eta^{(c)} e^{-\tau_{\ell}^{(c)}}\right) =  1 - \sum_{\ell=2}^{k-2} \exp\left(-\sum_{c \in C}\eta^{(c)} e^{-\tau_{\ell}^{(c)}}\right).\]
%
\noindent Note none of the above uses our boundedness assumptions on the birth or death rates $\lambda^{(c)}, \mu^{(c)}$. Using these assumptions on the birth and death rates, we see:
%
\[\P(\cB \subseteq \cX_1) \geq 1 - \sum_{\ell=2}^{k-2}q_\ell^{|C_N|} \geq 1-(k-3)q_{k-2}^{|C_N|}, \qquad q_\ell := \exp\left(-bT_{min} \exp\left[-2\ell DT_{max}\right]\right).\]
%
\noindent Hence we get the desired exponential convergence of the converge probability.
\end{proof}


%%%%%%%%%%%%%%%
% Sample complexity bounds for new ASTRAL procedure
%%%%%%%%%%%%%%%
\begin{proof}[Proof of Theorem~\ref{thm:sample_complexity_separate_characters} (Sample Complexity PCH-ASTRAL(sc))]
    Note for a specific leaf set $s$, we have $n_s \sim \mathrm{Binomial}(n, p_s)$ informative characters (characters where $I_s$ occurs).
    Applying Chernoff's bound
        \[
            \P\left(n_s < \frac{np_s}{2}\right)
            \leq \exp\left(-\frac{np_s}{8}\right)
            \leq \exp\left(-\frac{np_m}{8}\right).
        \]
    Union bounding over all \(\binom{k}{4}\) quartets gives
        \[
            \P\left(\min_{s\in {[k]\choose 4}} n_s < \frac{np_m}{2}\right)
            \leq
            \binom{k}{4}\exp\left(-\frac{np_m}{8}\right).
        \]
    Thus, provided
        \[
            np_m \geq 8\log\left(\frac{\binom{k}{4}}{\varepsilon}\right),
        \]
    we have, with probability at least \(1-\varepsilon\), at least \(n'= np_m/2\) informative characters for every quartet.
            
    From here, we mimic the proof of Theorem 2.1 in~\cite{shekhar2018_how_many_genes_are_enough}. First, if for every quartet $q\in\cQ$ the modal topology among
    $\{\tau_q^{(c_i)}\}_{i=1}^{n'}$ is the true topology $\cT_q^*$, then exact ASTRAL returns
    the true tree $\cT^*$. Hence it suffices to union bound over quartets. Let $E_q$ denote
    the event that the true topology is not modal for quartet $q$.

    Fix a quartet $q$, and let $t_1,t_2$ denote the two incorrect topologies. Set
    \[
        p_q := \P(\tau_q^{(c)}=\cT_q^*),
        \qquad
        r_{t_i} := \P(\tau_q^{(c)}=t_i),
    \]
    \noindent By the anomaly-free property,
    $p_q>r_{t_i}$. Let
    \[
        \Delta_q := p_q-r_q,
        \qquad
        \Delta := \min_{q\in\cQ}\Delta_q >0.
    \]
    For this fixed quartet, let
    \[
        (N_*,N_1,N_2)\sim \mathrm{Mult}(n';p_q,r_q,r_q)
    \]
    count the number of characters assigning the true topology and the two incorrect
    topologies.

    Fix any $w<\Delta_q/3$, and define
    \[
        A_q := \left\{\frac{N_*}{n'}\in(p_q-w,p_q+w)\right\},
    \]
    and
    \[
        B_{q,j}:=\left\{\frac{N_j}{n'}\in(r_q-w,r_q+w)\right\},
        \qquad j=1,2.
    \]
    If $A_q\cap(B_{q,1}\cup B_{q,2})$ occurs, then the true topology is modal.
    Hence, by a union bound and Hoeffding's inequality,
    \[
        \P(E_q)
        \leq
        \P(A_q^c)+\P(B_{q,1}^c)
        \leq
        4\exp(-2n'w^2).
    \]
    Taking $w$ arbitrarily close to $\Delta_q/3$ gives
    \[
        \P(E_q)
        \leq
        4\exp\left(-\frac{2n'\Delta_q^2}{9}\right)
        \leq
        4\exp\left(-\frac{2n'\Delta^2}{9}\right).
    \]

    \noindent Finally, union bounding over all $\binom{k}{4}$ quartets,
    \[
        \P(\widehat{\cT}\neq \cT^*)
        \leq
        4\binom{k}{4}
        \exp\left(-\frac{2n'\Delta^2}{9}\right).
    \]
    Thus, to recover $\cT^*$ with probability at least $1-\varepsilon$, it suffices that
    \[
        n'
        >
        \frac{9}{2\Delta^2}
        \log\left(
            \frac{4\binom{k}{4}}{\varepsilon}
        \right).
    \]
\end{proof}

%%%%%%%%%%%%%%%%%%%
% Expected Margin
%%%%%%%%%%%%%%%%%%%
\begin{proof}[Proof of Lemma \ref{lem:canby_expected_margin} (Expected support margin)]

Suppose first that the quartet $q = uv || xy$ induces a balanced tree in $\cT^*$. Let $i, j, k$ denote the internal vertices of this subtree like so:

\begin{center}
\begin{tikzpicture}[
  level distance=12mm,
  level 1/.style={sibling distance=24mm},
  level 2/.style={sibling distance=14mm},
  every node/.style={font=\small},
  edge from parent/.style={draw,-},
]
\node (i) {$i$}
  child { node (j) {$j$}
    child { node (u) {$u$} }
    child { node (v) {$v$} }
  }
  child { node (k) {$k$}
    child { node (x) {$x$} }
    child { node (y) {$y$} }
  };
\end{tikzpicture}
\end{center}

\noindent Let $I := c(i)$ be the set of micro-states at vertex $i$, and likewise for $J,K$. Then:
%
\[\E [\Delta_q^{(c)} \mid J, K] = \sum_{(\alpha, \beta) \in \cA(J, K)} \E(I_{uv||xy}(\alpha, \beta) \mid J, K) \]
%
\noindent where:
%
\[\cA(J, K) = \{(\alpha, \beta) : \alpha\neq\beta \in \Sigma_N, \alpha \in J , \beta \in K, \{\alpha, \beta\} \nsubseteq J \cap K\}\]
%
\noindent We see that:
%
\[ \cB(J, K) := \{(\alpha, \beta) : \alpha \in J \setminus K, \beta \in K \setminus J \}  \subseteq \cA(J, K)\]
%
\noindent For $(\alpha, \beta) \in \cB(J, K)$, since we just need $\alpha, \beta$ to survive until the leaves, by Lemma \ref{lem:death_rate_prob_bounds}: 
%
\[\E(I_{uv||xy}(\alpha, \beta) \mid J, K) \geq \eta_{ju} \eta_{jv} \eta_{kx} \eta_{ky}\]
%
\noindent Hence:
%
\[\E \Delta_q^{(c)} \geq \E |\cB(J, K)| \cdot\eta_{ju} \eta_{jv} \eta_{kx} \eta_{ky}\]
%
\noindent Note $(\alpha, \beta) \in \cB(J, K)$ can occur if:
%
\begin{enumerate}[label=(\roman*)]
    \item State $\alpha$ is born along the branch $(i, j]$ and survives until $j$. State $\beta$ is born along the branch $(i, k]$ and survives until $k$. 

    \item State $\alpha$ is born before $i$, dies along $(i,k]$, and survives along $(i, j]$. State $\beta$ is born along the branch $(i, k]$ and survives until $k$. 

    \item State $\beta$ is born before $i$, dies along $(i,j]$, and survives along $(i, k]$. State $\alpha$ is born along the branch $(i, j]$ and survives until $j$. 

    \item  State $\alpha$ is born before $i$, dies along $(i,k]$, and survives along $(i, j]$. State $\beta$ is born before $i$, dies along $(i,j]$, and survives along $(i, k]$.
\end{enumerate}
%
\noindent We focus solely on $(i)$, as in this case the roles of $J,K$ nicely decouple. By Lemma \ref{lem:pop_sizes_in_ID_processes}, the number of such $\alpha$ and $\beta$ 
are at least $\mathrm{Poisson(m_0(\ell_{ij}))}$ and $\mathrm{Poisson(m_0(\ell_{ik}))}$ 
respectively, where $m_\theta(t) = \theta e^{-\mu t} + \frac{\lambda_N}{\mu} \cdot (1-\eta_t)$, and these are 
clearly independent. This produces on average $m_0(\ell_{ij}) \cdot m_0(\ell_{ik})$ pairs $(\alpha, \beta)$. 

The contributions from (ii), (iii), (iv) can similarly be calculated using the death-rate lower bound
found in Lemma~\ref{lem:death_rate_prob_bounds}. If we let $\theta = \E|R \cap \Sigma_N|$ be the expected
number of non-homoplastic states at the root, then by Poisson thinning we get contributions
%
\begin{enumerate}
    \item[(ii)] $m_\theta(\ell_{ri}) \xi_{ik} \eta_{ij} \cdot m_0(\ell_{ik})$. 
    \item[(iii)] $m_0(\ell_{ij}) \cdot m_\theta(\ell_{ri}) \xi_{ij} \eta_{ik} $.
\end{enumerate}

\noindent If we additionally assume $|R \cap \Sigma_N| \sim \mathrm{Poisson}(\theta)$, then
\begin{enumerate}
        \item[(iv)] $m_\theta(\ell_{ri})^2 \xi_{ik} \eta_{ij} \xi_{ij} \eta_{ik}$. 
\end{enumerate}
%
\noindent Formula (iv) comes from the fact that $|I| \sim \mathrm{Poisson}(m_\theta(\ell_{ri}))$, and so there are $\E |I| (|I| - 1) = m_\theta(\ell_{ri})^2$ ordered pairs 
of distinct states on average at $i$. Hence when the subtree induced by the quartet is balanced:
%
\[\E |\cB(J,K)|
\geq
\left[
m_0(\ell_{ij})
+
m_\theta(\ell_{ri})\eta_{ij}\xi_{ik}
\right]
\left[
m_0(\ell_{ik})
+
m_\theta(\ell_{ri})\xi_{ij}\eta_{ik}
\right] \geq m_0(\ell_{ij})  m_0(\ell_{ik}).\]
%
\noindent where the right hand side is always a valid lower bound even without the Poisson
assumption on $|R|$. With our above work, this gives the desired bound on $\E \Delta_q^{(c)}$ when the induced subtree is balanced.

If the subtree induced by $q$ is a caterpillar tree like so:

\begin{center}
\begin{tikzpicture}[
  level distance=12mm,
  level 1/.style={sibling distance=24mm},
  level 2/.style={sibling distance=14mm},
  every node/.style={font=\small},
  edge from parent/.style={draw,-},
]
\node (i) {$i$}
  child { node (u) {$u$}}
  child { node (j) {$j$}
    child { node (v) {$v$} }
    child { node (k) {$k$} 
        child { node (x) {$x$} }
        child { node (y) {$y$} }
    }
  };
\end{tikzpicture}
\end{center}

\noindent Then we saw:
%
\[\E [\Delta_q^{(c)} \mid I, J, K] = \sum_{(\alpha, \beta) \in \tilde{\cA}(I, J, K)} \E(I_{uv||xy}(\alpha, \beta) \mid I, J, K) \]
%
\noindent where:
\[\tilde{\cA}(I, J, K) = \{(\alpha, \beta) : \alpha\neq\beta \in \Sigma_N, \alpha \in I \cap J , \beta \in K, \{\alpha, \beta\} \nsubseteq J \cap K\}\]
%
\noindent Again, we see this set contains:
%
\[ \tilde{\cB}(I, J, K) := \{(\alpha, \beta) : \alpha \in (I \cap J) \setminus K, \beta \in K \setminus J \} \]
%
For $(\alpha, \beta) \in \tilde{\cB}(I, J, K)$ we see $(\alpha,\beta) \in M_{uv||xy}$ if and only if:

\begin{itemize}
    \item $\alpha$ survives on $(i, u]$ and $(j, v]$.
    \item $\beta$ survives on $(k,x]$ and $(k,y]$
\end{itemize}

\noindent so again by Lemma \ref{lem:death_rate_prob_bounds}:
%
\[\E(I_{uv||xy}(\alpha, \beta) \mid I, J, K) \geq \eta_{iu}\eta_{jv} \eta_{kx} \eta_{ky}\]
%
\noindent Note $(\alpha, \beta) \in \tilde{B}(I,J,K)$ can occur only if $\alpha$ is born before $i$, survives along $(i, j]$ and dies on $(j,k]$,
and $\beta$ is born on $(j,k]$ and survives until $k$. Let $r$ denote the root of the tree. Note for $\alpha$, any 
non-homoplastic state born at the root survives to $i$ with probability at least $\eta_{ri}$. Independently from this, we have at 
least $\mathrm{Poisson(m_0(\ell_{ri})})$ born along the path $[r, i]$. Of these, at least a $\eta_{ij}\xi_{jk}$ fraction survive until they
 reach $j$ but die before reaching $k$. Independently of this, at least $\mathrm{Poisson(m_0(\ell_{jk}))}$ non-homoplastic states $\beta$ are born 
 along $(j,k]$ and do not die on this branch. Hence if we let $\rho_N := \E |c(r) \cap \Sigma_N|$ be the expected number of
  non-homoplastic states present at the root
%
\[\E |\tilde{B}(I,J,K)| \geq \eta_{ij} \xi_{jk}\left[\rho_N \cdot \eta_{ri} + \frac{\lambda_N}{\mu}(1-e^{-\mu \ell_{ri}})\right] \cdot \left[\frac{\lambda_N}{\mu} (1-e^{-\mu \ell_{jk}})\right],\]
%
\noindent giving the desired bound. 
\end{proof}
% END PROOFFFFFF




\begin{proof}[Proof of Lemma \ref{lem:sample_complexity_pch_astral} (Sample Complexity PCH-ASTRAL)]
    If for each leaf-set $s$ the topology $q \in A_s$ maximizing the score $S_q$ is the correct topology, 
    then PCH-ASTRAL will trivially return the correct species tree $\cT^*$, as the set of topologies $\{q_s\}_{s \in {[k] \choose 4}}$ 
    will exactly match those generated by the species tree $\cT^*$. Hence if $\hat{\cT}$ is the output of this inference procedure, by union bound:
    %
     \[\P(\hat{\cT} = \cT^*) \geq 1 - \sum_{s} \P(q_s \neq \cT^*_s).\]   
    %
    \noindent Hence it suffices to upper bound $p_s := \P(q_s \neq \cT^*_s)$. Note
    %
    \[p_s \leq \sum_{t \in A_s \setminus \{\cT_s^*\}} \P(S_{\cT_s^*} \leq S_t).\]
    %
    \noindent so it suffices to bound each term on the right hand side. For $t \in A_s \setminus \{\cT_s^*\}$, define
    %
        \[D_i := |M^{(c_i)}_{\cT^*_s}| - |M^{(c_i)}_t|, \qquad S_{\cT_s^*} - S_t = \frac{1}{n} \sum_{i=1}^n D_i. \]
    %
    \noindent Since $L \leq \E D_i$ we have
    %
        \[\P(S_{\cT_s^*} \leq S_t) \leq \P\left(\Bigg| \sum_{i=1}^n (D_i - \E D_i)\Bigg| \geq n L\right).\]
    %
    \noindent Lemma~\ref{lem:canby_concentration_of_scores} implies for universal constants
    %
        \[||D_i - \E D_i ||_{\Psi_{1/2}} \lesssim \Big||M_{\cT_s^*}| \Big|_{\Psi_{1/2}} + \Big||M_t| \Big|_{\Psi_{1/2}} \lesssim \kappa,\]
    %

    \noindent Applying the sub-Weibull concentration inequality of Theorem 3.1 in~\cite{kuchibhotla2022weibull}
    %
        \[\P\left(\Bigg| \sum_{i=1}^n (D_i - \E D_i)\Bigg| \gtrsim \kappa( \sqrt{nt} + t^2)\right) \leq 2e^{-t}.\]
    %
    \noindent Choose $t_\delta = \log(4{k \choose 4} / \delta)$. Note our assumptions on $n$ imply
    %
    \[nL \gtrsim \kappa (\sqrt{nt_\delta} + t_\delta^2)\]
    %
    \noindent and we have
    %
        \[\P(S_{\cT_s^*} \leq S_t) \leq 2\exp(-t_\delta).\]
    %
    \noindent Unioning-bounding over quartets and the two incorrect topologies then gives
    %
        \[\P(\hat{\cT} = \cT^*) \geq 1 - 2{k \choose 4} \cdot 2\exp(-t_\delta) = 1-\delta, \]
    %
    \noindent as desired.
\end{proof}

\end{subappendices}



